---
title: "Taxa"
author: "HP"
date: "10/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## library
```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
library(tidyverse)
library(phyloseq)
library(picante)
library(Hmisc)
library(agricolae)
library(vegan)
library(ape)
library(usedist)
library(grid)
library(gridExtra)
library(gtable)
library(ggpubr)
library(ggrepel)
library(readxl)
library(ComplexHeatmap)
```

## data filtering
```{r}
load(file = "temp.RData")

meta = meta_16S[pmatch(rownames(as.matrix(dm_bray)),meta_16S$SampleID,),]

meta_pup <- meta %>% filter(Generation == "F1") %>% 
    mutate(Facility = case_when(Facility == "Farm" ~ "Farm", 
                                T ~ "Vivarium"),
           Treatment = case_when(TREATMENT == "DIWATER" ~ "Diwater",
                                 T ~ "Chlorinated"),
           Sex = case_when(Sex == "F" ~ "Female",
                            T ~ "Male")) %>% 
    mutate(SamplingTimepoint = str_trim(SamplingTimepoint)) %>%
    mutate(Week = case_when(SamplingTimepoint == "Week3" ~ "Week 3",
                            SamplingTimepoint == "Week7" ~ "Week 7",
                            SamplingTimepoint == "Week12" ~ "Week 12",
                            T ~ NA)) %>% 
    drop_na(Week)

vivarium_IDs <- filter(meta_pup, Week == "Week 12" & Facility == "Vivarium") %>% distinct(SubjectID) %>% pull(SubjectID)

farm_IDs <- filter(meta_pup, Week == "Week 7" & Facility == "Farm") %>% 
  group_by(FamilyID) %>%
  filter(n() > 1) %>%
  ungroup() %>% 
  distinct(SubjectID) %>% pull(SubjectID)

final_IDs <- c(vivarium_IDs, farm_IDs)
meta_filtered <- meta_pup %>% filter(SubjectID %in% final_IDs)

theme_all <- theme(panel.background = element_rect(fill = NA), 
                   plot.title = element_text(hjust = 0.5), 
                   axis.text = element_text(size = 8),
                   axis.title = element_text(size = 10),
                   legend.text = element_text(size=10),
                   legend.title = element_blank(),
                   title = element_text(size=10),
                   strip.text = element_text(color = "black", face = "bold"),
                   strip.background = element_blank(),
                   aspect.ratio = 1
                   )

```

## taxa
### masslin2
#### function
```{r message=FALSE, warning=FALSE, include=FALSE}
library(Maaslin2)
### function wrap for all taxa levels
#in put metadata, sample id column name, phyloseq object
do_maaslin2 = function(meta, sample_id, phylo_rf, fixed, random, ref, output, 
                       min_prevalence = 0.1, min_abundance = 0.0001, max_significance = 0.05,
                       normalization = "TSS", transform = "LOG"){
  ft = otu_table(phylo_rf@otu_table) %>% as.matrix() %>% as.data.frame() %>% rownames_to_column(var = "id")
  wft = ft %>% select(id,  meta %>% pull(sample_id)) %>% column_to_rownames(var = "id") %>% t()
  wmeta = meta %>% column_to_rownames(var = sample_id)
  taxrank = "ASV"
  fit = Maaslin2(
    input_data = wft, 
    input_metadata = wmeta, 
    normalization = normalization,
    transform = transform,
    output = paste0(output,"_",taxrank),
    plot_heatmap = F,
    plot_scatter = F,
    fixed_effects = fixed,
    reference = ref,
    random_effects = random,
    min_prevalence = min_prevalence,
    min_abundance = min_abundance,
    max_significance = max_significance
  )
  for (taxrank in c("Species","Genus","Family","Order","Class","Phylum")){
  phylo_glom = tax_glom(phylo_rf, taxrank = taxrank, NArm = F)
  ft_glom = otu_table(phylo_glom@otu_table) %>% as.matrix() %>% as.data.frame() %>% rownames_to_column(var = "id")
  wft = ft_glom %>% select(id, meta %>% pull(sample_id)) %>% column_to_rownames(var = "id") %>% t()
  fit = Maaslin2(
      input_data = wft, 
      input_metadata = wmeta, 
      normalization = normalization,
      transform = transform,
      output = paste0(output,"_",taxrank),
      plot_heatmap = F,
      plot_scatter = F,
      fixed_effects = fixed,
      reference = ref,
      random_effects = random,
      min_prevalence = min_prevalence,
      min_abundance = min_abundance,
      max_significance = max_significance
      )
  }
}

summarise_maaslin2 = function(path, group, taxa){
  res = matrix(data = NA, nrow = 1, ncol = 20)
  res = as.data.frame(res)
  names(res) =  c("feature","metadata","value","coef","stderr","N","N.not.0","pval","qval","Group",
                  "id","Kingdom","Phylum","Class","Order","Family","Genus","Species","ASV", "level")
  for (taxrank in c("ASV", "Species", "Genus", "Family", "Order", "Class", "Phylum")){
    a = read.table(file = paste0(path ,"_", taxrank, "/significant_results.tsv"),
                  header = T, sep = "\t", stringsAsFactors = F)
    if (nrow(a) == 0) {next}
    a = a %>% mutate(Group = group, id = str_remove(feature,"^X"),feature = str_remove(feature,"^X")) %>% left_join(taxa, by = "id") %>% mutate(level = taxrank)
    res = rbind(res,a)
  }
  res = res %>% filter(!is.na(feature))
  res
}

```

#### compare
```{r message=FALSE, warning=FALSE, include=FALSE}
output = "output/taxa/maaslin2/"
#meta_all = phylo_rf@sam_data@.Data %>% as.data.frame()
#names(meta_all) = phylo_rf@sam_data@names
#meta_all$SampleID = phylo_rf@sam_data@row.names

### All_moms
meta = meta_all %>%  filter(Generation == "F0") %>% mutate(SamplingTimepoint_mom = as.character(SamplingTimepoint_mom))
#### include all timepoint
fit = do_maaslin2(meta = meta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                  fixed = c("TREATMENT", "Facility", "SamplingTimepoint_mom"), 
                  random = c("SubjectID"), 
                  ref = c("TREATMENT,DIWATER", "Facility,Bartlett", "SamplingTimepoint_mom,E16"), 
                  output = paste0(output,"All_moms")
                  )
res = summarise_maaslin2(path = "output/taxa/maaslin2/All_moms", group = "Mom_compare_all", taxa)
write.csv(res, file = "output/taxa/maaslin2/summary_All_moms.csv",quote = F, row.names = F)
#### All_moms by time points
for (week in meta$SamplingTimepoint_mom %>% unique()){
  wmeta = meta %>% filter(SamplingTimepoint_mom == week)
  fit = do_maaslin2(meta = wmeta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                  fixed = c("TREATMENT", "Facility"), 
                  random = NULL, 
                  ref = c("TREATMENT,DIWATER", "Facility,Bartlett"), 
                  output = paste0(output,"moms_",week)
                  )
  res = summarise_maaslin2(path = paste0(output,"moms_",week), group = week, taxa)
  write.csv(res, file = paste0("output/taxa/maaslin2/summary_moms_",week, ".csv"),quote = F, row.names = F)  
} 

### All_pups
meta = meta_all %>%  filter(Generation == "F1")
for (s in c("M", "F")){
  wmeta = meta %>% filter(Sex == s)
  fit = do_maaslin2(meta = wmeta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                  fixed = c("TREATMENT", "Facility", "SamplingTimepoint"), 
                  random = NULL, 
                  ref = c("TREATMENT,DIWATER", "Facility,Bartlett", "SamplingTimepoint,Week3"), 
                  output = paste0(output,"pups_",s)
                  )
  res = summarise_maaslin2(path = paste0(output,"pups_",s), group = s, taxa)
  write.csv(res, file = paste0("output/taxa/maaslin2/summary_pups_",s, ".csv"),quote = F, row.names = F)   
  for (week in c(meta$SamplingTimepoint %>% unique())){
    wmeta = meta %>% filter(Sex == s, SamplingTimepoint == week)
    fit = do_maaslin2(meta = wmeta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                    fixed = c("TREATMENT", "Facility"), 
                    random = NULL, 
                    ref = c("TREATMENT,DIWATER", "Facility,Bartlett"), 
                    output = paste0(output,"pups_",s, "_", week)
                    )
    res = summarise_maaslin2(path = paste0(output,"pups_",s, "_", week), group = paste0(s, "_", week), taxa)
    write.csv(res, file = paste0("output/taxa/maaslin2/summary_pups_",s, "_", week, ".csv"),quote = F, row.names = F)  
  }
}

### DI water compare facility
meta = meta_all %>%  filter(Generation == "F1", TREATMENT == "DIWATER")
for (s in c("M", "F")){
  wmeta = meta %>% filter(Sex == s)
  fit = do_maaslin2(meta = wmeta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                  fixed = c("Facility", "SamplingTimepoint"), 
                  random = NULL, 
                  ref = c("Facility,Bartlett", "SamplingTimepoint,Week3"), 
                  output = paste0(output,"pups_in_DIWATER_",s)
                  )
  res = summarise_maaslin2(path = paste0(output,"pups_in_DIWATER_",s), group = s, taxa)
  write.csv(res, file = paste0("output/taxa/maaslin2/summary_pups_in_DIWATER_",s, ".csv"),quote = F, row.names = F)   
  for (week in c(meta$SamplingTimepoint %>% unique())){
    wmeta = meta %>% filter(Sex == s, SamplingTimepoint == week)
    fit = do_maaslin2(meta = wmeta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                    fixed = c("Facility"), 
                    random = NULL, 
                    ref = c("Facility,Bartlett"), 
                    output = paste0(output,"pups_in_DIWATER_",s, "_", week)
                    )
    res = summarise_maaslin2(path = paste0(output,"pups_in_DIWATER_",s, "_", week), group = paste0(s, "_", week), taxa)
    write.csv(res, file = paste0("output/taxa/maaslin2/summary_pups_in_DIWATER_",s, "_", week, ".csv"),quote = F, row.names = F)  
  }
}

### CHLORINATED water compare facility
meta = meta_all %>%  filter(Generation == "F1", TREATMENT == "CHLORINATED")
for (s in c("M", "F")){
  wmeta = meta %>% filter(Sex == s)
  fit = do_maaslin2(meta = wmeta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                  fixed = c("Facility", "SamplingTimepoint"), 
                  random = NULL, 
                  ref = c("Facility,Bartlett", "SamplingTimepoint,Week3"), 
                  output = paste0(output,"pups_in_CHLORINATED_",s)
                  )
  res = summarise_maaslin2(path = paste0(output,"pups_in_CHLORINATED_",s), group = s, taxa)
  write.csv(res, file = paste0("output/taxa/maaslin2/summary_pups_in_CHLORINATED_",s, ".csv"),quote = F, row.names = F)   
  for (week in c(meta$SamplingTimepoint %>% unique())){
    wmeta = meta %>% filter(Sex == s, SamplingTimepoint == week)
    fit = do_maaslin2(meta = wmeta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                    fixed = c("Facility"), 
                    random = NULL, 
                    ref = c("Facility,Bartlett"), 
                    output = paste0(output,"pups_in_CHLORINATED_",s, "_", week)
                    )
    res = summarise_maaslin2(path = paste0(output,"pups_in_CHLORINATED_",s, "_", week), group = paste0(s, "_", week), taxa)
    write.csv(res, file = paste0("output/taxa/maaslin2/summary_pups_in_CHLORINATED_",s, "_", week, ".csv"),quote = F, row.names = F)  
  }
}

### Bartlett compare treatment
meta = meta_all %>%  filter(Generation == "F1", Facility == "Bartlett")
for (s in c("M", "F")){
  wmeta = meta %>% filter(Sex == s)
  fit = do_maaslin2(meta = wmeta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                  fixed = c("TREATMENT", "SamplingTimepoint"), 
                  random = NULL, 
                  ref = c("TREATMENT,DIWATER","SamplingTimepoint,Week3"), 
                  output = paste0(output,"pups_in_Bartlett_",s)
                  )
  res = summarise_maaslin2(path = paste0(output,"pups_in_Bartlett_",s), group = s, taxa)
  write.csv(res, file = paste0("output/taxa/maaslin2/summary_pups_in_Bartlett_",s, ".csv"),quote = F, row.names = F)   
  for (week in c(meta$SamplingTimepoint %>% unique())){
    wmeta = meta %>% filter(Sex == s, SamplingTimepoint == week)
    fit = do_maaslin2(meta = wmeta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                    fixed = c("TREATMENT"), 
                    random = NULL, 
                    ref = c("TREATMENT,DIWATER"), 
                    output = paste0(output,"pups_in_Bartlett_",s, "_", week)
                    )
    res = summarise_maaslin2(path = paste0(output,"pups_in_Bartlett_",s, "_", week), group = paste0(s, "_", week), taxa)
    write.csv(res, file = paste0("output/taxa/maaslin2/summary_pups_in_Bartlett_",s, "_", week, ".csv"),quote = F, row.names = F)  
  }
}

### Farm compare treatment
meta = meta_all %>%  filter(Generation == "F1", Facility == "Farm")
for (s in c("M", "F")){
  wmeta = meta %>% filter(Sex == s)
  fit = do_maaslin2(meta = wmeta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                  fixed = c("TREATMENT", "SamplingTimepoint"), 
                  random = NULL, 
                  ref = c("TREATMENT,DIWATER","SamplingTimepoint,Week3"), 
                  output = paste0(output,"pups_in_Farm_",s)
                  )
  res = summarise_maaslin2(path = paste0(output,"pups_in_Farm_",s), group = s, taxa)
  write.csv(res, file = paste0("output/taxa/maaslin2/summary_pups_in_Farm_",s, ".csv"),quote = F, row.names = F)   
  for (week in c(meta$SamplingTimepoint %>% unique())){
    wmeta = meta %>% filter(Sex == s, SamplingTimepoint == week)
    fit = do_maaslin2(meta = wmeta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                    fixed = c("TREATMENT"), 
                    random = NULL, 
                    ref = c("TREATMENT,DIWATER"), 
                    output = paste0(output,"pups_in_Farm_",s, "_", week)
                    )
    res = summarise_maaslin2(path = paste0(output,"pups_in_Farm_",s, "_", week), group = paste0(s, "_", week), taxa)
    write.csv(res, file = paste0("output/taxa/maaslin2/summary_pups_in_Farm_",s, "_", week, ".csv"),quote = F, row.names = F)  
  }
}



```

### ancom
#### function
```{r include=FALSE}
library(nlme)
library(tidyverse)
library(ggplot2)
library(compositions)
source("ANCOM-master/scripts/ancom_v2.1.R")
### function wrap
do_ancom = function(meta_data, phylo_rf, group = "Mom_treatment", output,
                    group_var = "TREATMENT", sample_var = "SampleID", 
                    out_cut = 0.05, zero_cut = 0.9, lib_cut = 1000, neg_lb = FALSE,
                    main_var = "TREATMENT", p_adj_method = "BH", alpha = 0.05,
                    adj_formula = "Facility*SamplingTimepoint", rand_formula = NULL){
  taxrank = "ASV"
  feature_table = otu_table(phylo_rf@otu_table) %>% as.matrix() %>% as.data.frame() %>% 
      rownames_to_column(var = "id") %>% select(id, meta_data$SampleID) %>% column_to_rownames(var = "id")
  prepro = feature_table_pre_process(feature_table = feature_table,
                                     meta_data = meta_data,
                                     sample_var = sample_var,
                                     group_var = group_var,
                                     out_cut = out_cut,
                                     zero_cut = zero_cut,
                                     lib_cut = lib_cut,
                                     neg_lb = neg_lb)
  feature_table = prepro$feature_table
  meta_data = prepro$meta_data
  struc_zero = prepro$structure_zeros
  res = ANCOM(feature_table, meta_data, struc_zero, 
                main_var = main_var, 
                p_adj_method = p_adj_method, alpha = alpha, adj_formula = adj_formula, rand_formula = rand_formula)
  
  out = res$out %>% arrange(desc(W)) %>% filter(detected_0.6) %>% 
    left_join(res$fig$data,by = "taxa_id") %>% select(-y) %>% rename(CLR_diff = x)
  write.csv(out, file = paste0(output, group, "_", taxrank, "_res.csv"), row.names = F)

  for (taxrank in c("Species","Genus","Family","Order","Class","Phylum")){
    phylo_glom = tax_glom(phylo_rf, taxrank = taxrank, NArm = F)
    ft_glom = otu_table(phylo_glom@otu_table) %>% as.matrix() %>% as.data.frame() %>% rownames_to_column(var = "id")
    feature_table = ft_glom %>% select(id, meta_data$SampleID) %>% column_to_rownames(var = "id")
    prepro = feature_table_pre_process(feature_table = feature_table,
                                     meta_data = meta_data,
                                     sample_var = sample_var,
                                     group_var = group_var,
                                     out_cut = out_cut,
                                     zero_cut = zero_cut,
                                     lib_cut = lib_cut,
                                     neg_lb = neg_lb)
    feature_table = prepro$feature_table
    meta_data = prepro$meta_data
    struc_zero = prepro$structure_zeros
    res = ANCOM(feature_table, meta_data, struc_zero, 
                main_var = main_var, 
                p_adj_method = p_adj_method, alpha = alpha, adj_formula = adj_formula, rand_formula = rand_formula)
    out = res$out %>% arrange(desc(W)) %>% filter(detected_0.6) %>% 
      left_join(res$fig$data,by = "taxa_id") %>% select(-y) %>% rename(CLR_diff = x)
    write.csv(out, file = paste0(output, group, "_", taxrank, "_res.csv"), row.names = F)
  }
}

summarise_ancom = function(path, group, taxa){
  res = matrix(data = NA, nrow = 1, ncol = 19)
  res = as.data.frame(res)
  names(res) =  c("taxa_id","W","detected_0.9","detected_0.8","detected_0.7","detected_0.6","CLR_diff","zero_ind",
                  "Group", "id","Kingdom","Phylum","Class","Order","Family","Genus","Species","ASV", "level")
  for (taxrank in c("ASV", "Species", "Genus", "Family", "Order", "Class", "Phylum")){
    a = read.csv(file = paste0(path, group,"_", taxrank, "_res.csv"), stringsAsFactors = F)
    if (nrow(a) == 0) {next}
    a = a %>% mutate(Group = group, id = taxa_id) %>% left_join(taxa, by ="id") %>% mutate(level = taxrank)
    res = rbind(res,a)
  }
  res = res %>% filter(!is.na(id))
  res
}
```


#### compare
```{r include=FALSE}
output = "output/taxa/ancom/"
meta_all = phylo_rf@sam_data@.Data %>% as.data.frame()
names(meta_all) = phylo_rf@sam_data@names
meta_all$SampleID = phylo_rf@sam_data@row.names

### All_moms
meta_data = meta_all %>% filter(Generation == "F0")
fit = do_ancom(meta_data, phylo_rf, group = "Mom_TREATMENT", output,
                    group_var = "TREATMENT", sample_var = "SampleID", 
                    out_cut = 0.05, zero_cut = 0.9, lib_cut = 1000, neg_lb = FALSE,
                    main_var = "TREATMENT", p_adj_method = "BH", alpha = 0.05,
                    adj_formula = "Facility+SamplingTimepoint_mom", rand_formula = NULL)
res = summarise_ancom(path = "output/taxa/ancom/", group = "Mom_TREATMENT", taxa)
write.csv(res, file = "output/taxa/ancom/summary_Mom_TREATMENT.csv",quote = F, row.names = F)

fit = do_ancom(meta_data, phylo_rf, group = "Mom_Facility", output,
                    group_var = "Facility", sample_var = "SampleID", 
                    out_cut = 0.05, zero_cut = 0.9, lib_cut = 1000, neg_lb = FALSE,
                    main_var = "Facility", p_adj_method = "BH", alpha = 0.05,
                    adj_formula = "TREATMENT+SamplingTimepoint_mom", rand_formula = NULL)
res = summarise_ancom(path = "output/taxa/ancom/", group = "Mom_Facility", taxa)
write.csv(res, file = "output/taxa/ancom/summary_Mom_Facility.csv",quote = F, row.names = F)

for (week in meta_data$SamplingTimepoint_mom %>% unique()){
  wmeta = meta_all %>% filter(Generation == "F0", SamplingTimepoint_mom == week)
  fit = do_ancom(wmeta, phylo_rf, group = paste0("Mom_TREATMENT","_",week), output,
                    group_var = "TREATMENT", sample_var = "SampleID", 
                    out_cut = 0.05, zero_cut = 0.9, lib_cut = 1000, neg_lb = FALSE,
                    main_var = "TREATMENT", p_adj_method = "BH", alpha = 0.05,
                    adj_formula = "Facility", rand_formula = NULL)
  res = summarise_ancom(path = "output/taxa/ancom/", group = paste0("Mom_TREATMENT","_",week), taxa)
  write.csv(res, file = paste0("output/taxa/ancom/summary_Mom_TREATMENT_", week,".csv"),quote = F, row.names = F)
  
  fit = do_ancom(wmeta, phylo_rf, group = paste0("Mom_Facility","_",week), output,
                      group_var = "Facility", sample_var = "SampleID", 
                      out_cut = 0.05, zero_cut = 0.9, lib_cut = 1000, neg_lb = FALSE,
                      main_var = "Facility", p_adj_method = "BH", alpha = 0.05,
                      adj_formula = "TREATMENT", rand_formula = NULL)
  res = summarise_ancom(path = "output/taxa/ancom/", group = paste0("Mom_Facility","_",week), taxa)
  write.csv(res, file = paste0("output/taxa/ancom/summary_Mom_Facility_", week, ".csv"),quote = F, row.names = F)
}

### All_pups
meta_data = meta_all %>% filter(Generation == "F1")
for (s in c("M", "F")){
  wmeta = meta_data %>% filter(Sex == s)
  fit = do_ancom(wmeta, phylo_rf, group = paste0("pup_TREATMENT_",s), output,
                    group_var = "TREATMENT", sample_var = "SampleID", 
                    out_cut = 0.05, zero_cut = 0.9, lib_cut = 1000, neg_lb = FALSE,
                    main_var = "TREATMENT", p_adj_method = "BH", alpha = 0.05,
                    adj_formula = "Facility+SamplingTimepoint", rand_formula = NULL)
  res = summarise_ancom(path = "output/taxa/ancom/", group = paste0("pup_TREATMENT","_",s), taxa)
  write.csv(res, file = paste0("output/taxa/ancom/summary_pup_TREATMENT_", s, ".csv"),quote = F, row.names = F)
  
  fit = do_ancom(wmeta, phylo_rf, group =  paste0("pup_Facility_",s), output,
                      group_var = "Facility", sample_var = "SampleID", 
                      out_cut = 0.05, zero_cut = 0.9, lib_cut = 1000, neg_lb = FALSE,
                      main_var = "Facility", p_adj_method = "BH", alpha = 0.05,
                      adj_formula = "TREATMENT+SamplingTimepoint", rand_formula = NULL)
  res = summarise_ancom(path = "output/taxa/ancom/", group = paste0("pup_Facility","_",s), taxa)
  write.csv(res, file = paste0("output/taxa/ancom/summary_pup_Facility_", s, ".csv"),quote = F, row.names = F)    

  for (week in wmeta$SamplingTimepoint %>% unique()){
    wwmeta = wmeta %>% filter(SamplingTimepoint == week)
    fit = do_ancom(wwmeta, phylo_rf, group = paste0("pup_TREATMENT_",s, "_",week), output,
                      group_var = "TREATMENT", sample_var = "SampleID", 
                      out_cut = 0.05, zero_cut = 0.9, lib_cut = 1000, neg_lb = FALSE,
                      main_var = "TREATMENT", p_adj_method = "BH", alpha = 0.05,
                      adj_formula = "Facility", rand_formula = NULL)
    res = summarise_ancom(path = "output/taxa/ancom/", group = paste0("pup_TREATMENT_",s, "_",week), taxa)
    write.csv(res, file = paste0("output/taxa/ancom/summary_pup_TREATMENT_", s,"_", week,".csv"),quote = F, row.names = F)
    
    fit = do_ancom(wwmeta, phylo_rf, group = paste0("pup_Facility_",s, "_",week), output,
                        group_var = "Facility", sample_var = "SampleID", 
                        out_cut = 0.05, zero_cut = 0.9, lib_cut = 1000, neg_lb = FALSE,
                        main_var = "Facility", p_adj_method = "BH", alpha = 0.05,
                        adj_formula = "TREATMENT", rand_formula = NULL)
    res = summarise_ancom(path = "output/taxa/ancom/", group = paste0("pup_Facility_",s, "_",week), taxa)
    write.csv(res, file = paste0("output/taxa/ancom/summary_pup_Facility_",s,"_", week, ".csv"),quote = F, row.names = F)
    }  
}

```

### ALDEX2
#### function
```{r}
library(ALDEx2)

do_aldex2 = function(meta_data, phylo_rf, fo = NULL, 
                    cond = NULL, test = "glm", denom = "all", 
                    output, group){
  taxrank = "ASV"
  feature_table = otu_table(phylo_rf@otu_table) %>% as.matrix() %>% as.data.frame() %>% 
    rownames_to_column(var = "id") %>% dplyr::select(id, meta_data$SampleID) %>% 
    column_to_rownames(var = "id")
  mm = model.matrix(as.formula(fo), meta_data)
  if(!is.null(fo)){
    cond = model.matrix(as.formula(fo), meta_data)
  }
  x.all = aldex(feature_table, conditions = mm, test = "glm", denom = "all")
  out = x.all %>% dplyr::select(starts_with("model")) %>% rownames_to_column(var = "id") %>%
    pivot_longer(-id, names_to = c("model", "var"), names_pattern = "model\\.(.*?)\\.(.*)") %>% 
    mutate(var = str_remove_all(var, "\\.")) %>% 
    pivot_wider(id_cols = c(id, model), names_from = "var", values_from = "value")
  write.csv(out, file = paste0(output, group, "_", taxrank, "_res.csv"), row.names = F)
  gc(verbose = F)
  for (taxrank in c("Species","Genus","Family","Order","Class","Phylum")){
      phylo_glom = tax_glom(phylo_rf, taxrank = taxrank, NArm = F)
      feature_table = otu_table(phylo_glom@otu_table) %>% as.matrix() %>% as.data.frame() %>% 
        rownames_to_column(var = "id") %>% dplyr::select(id, meta_data$SampleID) %>% 
        column_to_rownames(var = "id")
      x.all = aldex(feature_table, conditions = mm, test = "glm", denom = "all")
      out = x.all %>% dplyr::select(starts_with("model")) %>% rownames_to_column(var = "id") %>%
        pivot_longer(-id, names_to = c("model", "var"), names_pattern = "model\\.(.*?)\\.(.*)") %>% 
        mutate(var = str_remove_all(var, "\\.")) %>% 
        pivot_wider(id_cols = c(id, model), names_from = "var", values_from = "value")
      write.csv(out, file = paste0(output, group, "_", taxrank, "_res.csv"), row.names = F)
      gc(verbose = F)
  }
}

summarise_aldex2 = function(path, group, taxa){
  res = matrix(data = NA, nrow = 1, ncol = 17)
  res = as.data.frame(res)
  names(res) =  c("id","model","Estimate","StdError","tvalue","Prt","PrtBH","Group",
                  "Kingdom","Phylum","Class","Order","Family","Genus","Species","ASV", "level")
  for (taxrank in c("ASV", "Species", "Genus", "Family", "Order", "Class", "Phylum")){
    a = read.csv(file = paste0(path,group, "_", taxrank, "_res.csv"), stringsAsFactors = F)
    a = a %>% filter(PrtBH < 0.1) %>% mutate(Group = group) %>% 
      left_join(taxa, by = "id") %>% mutate(level = taxrank)
    res = rbind(res,a)
  }
  res %>% filter(!is.na(id))
}

```

#### compare
```{r}
output = "output/taxa/aldex2/"
meta_all = phylo_rf@sam_data@.Data %>% as.data.frame()
names(meta_all) = phylo_rf@sam_data@names
meta_all$SampleID = phylo_rf@sam_data@row.names

### All_moms
meta_data = meta_all %>% filter(Generation == "F0") %>% 
  mutate(SamplingTimepoint_mom = as.character(SamplingTimepoint_mom))

fit = do_aldex2(meta_data, phylo_rf, fo = "~ TREATMENT + Facility + SamplingTimepoint_mom",
               output = output, group = "mom")
res = summarise_aldex2(path = output, group = "mom", taxa)
write.csv(res, file = paste0("output/taxa/aldex2/summary_Mom_all", ".csv"), quote = F, row.names = F)

for (week in meta_data$SamplingTimepoint_mom %>% unique()){
  wmeta = meta_data %>% filter(SamplingTimepoint_mom == week)
  fit = do_aldex2(wmeta, phylo_rf, fo = "~ TREATMENT + Facility",
                 output = output, group = paste0("mom_", week))
  res = summarise_aldex2(path = output, group = paste0("mom_", week), taxa)
  write.csv(res, file = paste0("output/taxa/aldex2/summary_Mom_",week, ".csv"), quote = F, row.names = F)
}

### All_pups
meta_data = meta_all %>% filter(Generation == "F1") 

for (s in c("M", "F")){
  wmeta = meta_data %>% filter(Sex == s)
  fit = do_aldex2(meta_data, phylo_rf, fo = "~ TREATMENT + Facility + SamplingTimepoint",
                 output = output, group = paste0("pup_", s))
  res = summarise_aldex2(path = output, group = paste0("pup_", s), taxa)
  write.csv(res, file = paste0("output/taxa/aldex2/summary_Pup_",s, ".csv"), quote = F, row.names = F)
  
  for (week in meta_data$SamplingTimepoint %>% unique()){
    wmeta = meta_data %>% filter(Sex == s,SamplingTimepoint == week)
    fit = do_aldex2(wmeta, phylo_rf, fo = "~ TREATMENT + Facility",
                   output = output, group = paste0("pup_",s, "_", week))
    res = summarise_aldex2(path = output, group = paste0("pup_",s, "_", week), taxa)
    write.csv(res, file = paste0("output/taxa/aldex2/summary_Pup_", s, "_",week, ".csv"), quote = F, row.names = F)
  }
}
  
```

### compare different methods
```{r}
# with week 12 pups M facility
# masslin2
a = read.csv("output/taxa/maaslin2/summary_pups_M_Week12.csv")
picked_maaslin2 = a %>% filter(metadata == "Facility", level == "ASV") %>% pull(feature)
# ancom2
a = read.csv("output/taxa/ancom/pup_Facility_M_Week12_ASV_res.csv")
picked_ancom2 = a %>% pull(taxa_id) 
# aldex2
a = read.csv("output/taxa/aldex2/pup_M_Week12_ASV_res.csv")
picked_aldex2 = a %>% filter(str_detect(model, "Facility"), PrtBH < 0.1) %>% pull(id)
# venn
library(gplots)

a = venn(list("aldex"=picked_aldex2, "ancom"=picked_ancom2, "maaslin"=picked_maaslin2), 
         show.plot = F, intersections = F)
a
# valcano plot
log2FC_lm_estimate = function(wft, wmeta, fo, Var){
  # estimate log2 fold change using log10 transformed RA with simplified lm model
  coef_p_l = split(wft, rownames(wft)) %>% lapply(., function(x){
      log_RA = t(x)
      colnames(log_RA) = "log_RA"
      coefd = (merge(log_RA, wmeta, by.x = 0, by.y = "SampleID") %>%
                 lm(formula = as.formula(fo), data = .) %>%
                 summary())$coefficients
      coefd[which(str_detect(rownames(coefd), Var)),c("Estimate", "Pr(>|t|)")]
      })
  coef_and_p = do.call(rbind.data.frame, coef_p_l)
  names(coef_and_p) = c("coef","pvalue")
  rownames(coef_and_p) = names(coef_p_l)
  coef_and_p$padj = p.adjust(coef_and_p$pvalue, method = "BH")
  coef_and_p$Log2FC = log2(10^coef_and_p$coef)
  coef_and_p
}

all_asv = c(picked_aldex2,picked_ancom2,picked_maaslin2) %>% unique()
wmeta = meta_16S %>% filter(Generation == "F1", Sex == "M", SamplingTimepoint == "Week12")
ft = otu_table(phylo_rf@otu_table) %>% as.matrix() %>% as.data.frame() 
wft = ft[all_asv, wmeta %>% pull(SampleID)]
wft = log10((wft + 1)/colSums(ft)[1])
fo = "log_RA ~ TREATMENT + Facility "
Var = "Facility"

coef_and_p = log2FC_lm_estimate(wft, wmeta,
                       fo = "log_RA ~ TREATMENT + Facility",
                       Var = "TREATMENT")

coef_and_p = coef_and_p %>% rownames_to_column(var = "taxa") %>% 
  mutate(aldex = ifelse(taxa %in% picked_aldex2, "Y", "N"),
         ancom = ifelse(taxa %in% picked_ancom2, "Y", "N"),
         maaslin = ifelse(taxa %in% picked_maaslin2, "Y", "N"))
  
p1 = ggplot(data = coef_and_p, aes(x=Log2FC, y = -log10(pvalue))) + 
  geom_point(aes(color = aldex)) + 
  geom_vline(xintercept=c(-1,-0.5, 0.5,  1), col="red") + 
  geom_hline(yintercept=-log10(0.05), col="red") + 
  scale_color_manual(values = c("gray","red"))

p2 = ggplot(data = coef_and_p, aes(x=Log2FC, y = -log10(pvalue))) + 
  geom_point(aes(color = ancom)) + 
  geom_vline(xintercept=c(-1,-0.5, 0.5,  1), col="red") + 
  geom_hline(yintercept=-log10(0.05), col="red") + 
  scale_color_manual(values = c("gray","red"))

p3 = ggplot(data = coef_and_p, aes(x=Log2FC, y = -log10(pvalue))) + 
  geom_point(aes(color = maaslin)) + 
  geom_vline(xintercept=c(-1,-0.5, 0.5,  1), col="red") + 
  geom_hline(yintercept=-log10(0.05), col="red") +
  scale_color_manual(values = c("gray","red"))


```








### heatmaps
```{r message=FALSE, warning=FALSE, include=FALSE}
library(ComplexHeatmap)
### function wrap
log2FC_lm_estimate = function(wft, wmeta,taxa_id, fo, Var){
  # estimate log2 fold change using log10 transformed RA with simplified lm model
  coef_p_l = split(wft, wft[taxa_id]) %>% lapply(., function(x){
      log_RA = t(x[,-1])
      colnames(log_RA) = "log_RA"
      coefd = (merge(log_RA, wmeta, by.x = 0, by.y = "SampleID") %>%
                 lm(formula = as.formula(fo), data = .) %>%
                 summary())$coefficients
      coefd[which(str_detect(rownames(coefd), Var)),c("Estimate", "Pr(>|t|)")]
      })
  coef_and_p = do.call(rbind.data.frame, coef_p_l)
  names(coef_and_p) = c("coef","pvalue")
  rownames(coef_and_p) = names(coef_p_l)
  coef_and_p$padj = p.adjust(coef_and_p$pvalue, method = "BH")
  coef_and_p$Log2FC = log2(10^coef_and_p$coef)
  coef_and_p
}

mat_col_fun = circlize::colorRamp2(c(-4, -3.3, -2.6, -1.9, -1.2, -0.5), c("#253494","#2c7fb8","#41b6c4", "#7fcdbb", "#c7e9b4", "#ffffcc"))
col_fun = circlize::colorRamp2(c(-2, 0, 2), c("green", "white", "red"))
```


#### compare maaslin and ancom difference
```{r message=FALSE, warning=FALSE, include=FALSE}
#### ancom
a = lapply(c("Week3","Week4","Week5","Week7","Week9","Week12"),
           function(x){read.csv(paste0("output/taxa/ancom/summary_pups_TREATMENT_",x, ".csv"))})  
res_a_pup_TREATMENT = do.call(rbind, a) 
res_a_pup_TREATMENT_asv = res_a_pup_TREATMENT %>% filter(level == "ASV") %>% select(taxa_id, CLR_diff, Group) %>% 
  pivot_wider(names_from = "Group", values_from = "CLR_diff")
res_a_pup_TREATMENT_g = res_a_pup_TREATMENT %>% filter(level == "Genus") %>% select(taxa_id, CLR_diff, Group) %>% 
  pivot_wider(names_from = "Group", values_from = "CLR_diff")

a = lapply(c("Week3","Week4","Week5","Week7","Week9","Week12"),
           function(x){read.csv(paste0("output/taxa/ancom/summary_pups_Facility_",x, ".csv"))})  
res_a_pup_Facility = do.call(rbind, a)  
res_a_pup_Facility_asv = res_a_pup_Facility %>% filter(level == "ASV") %>% select(taxa_id, CLR_diff, Group) %>% 
  pivot_wider(names_from = "Group", values_from = "CLR_diff")
res_a_pup_Facility_g = res_a_pup_Facility %>% filter(level == "Genus") %>% select(taxa_id, CLR_diff, Group) %>% 
  pivot_wider(names_from = "Group", values_from = "CLR_diff")

#### maaslin
a = lapply(c("Week3","Week4","Week5","Week7","Week9","Week12"),
           function(x){read.csv(paste0("/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/output/taxa/maaslin2/summary_pups_",x, ".csv"))})  
res_m_pup_sum = do.call(rbind, a)  

res_m_pup_TREATMENT_asv = res_m_pup_sum %>% filter(metadata == "TREATMENT",level == "ASV") %>% select(feature, coef, Group) %>% pivot_wider(names_from = "Group", values_from = "coef")
res_m_pup_TREATMENT_g = res_m_pup_sum %>% filter(metadata == "TREATMENT",level == "Genus") %>% select(feature, coef, Group) %>% pivot_wider(names_from = "Group", values_from = "coef")

res_m_pup_Facility_asv = res_m_pup_sum %>% filter(metadata == "Facility",level == "ASV") %>% select(feature, coef, Group) %>% pivot_wider(names_from = "Group", values_from = "coef")
res_m_pup_Facility_g = res_m_pup_sum %>% filter(metadata == "Facility",level == "Genus") %>% select(feature, coef, Group) %>% pivot_wider(names_from = "Group", values_from = "coef")

#### diff
a = gplots::venn(list(m = res_m_pup_Facility_asv %>% pull(feature), a = res_a_pup_Facility_asv %>% pull(taxa_id)), show.plot = F)
attr(a, "class") = NULL
attr(a, "intersections") = NULL
a

a = gplots::venn(list(m = res_m_pup_TREATMENT_asv %>% pull(feature), a = res_a_pup_TREATMENT_asv %>% pull(taxa_id)), show.plot = F)
attr(a, "class") = NULL
attr(a, "intersections") = NULL
a

a = gplots::venn(list(m = res_m_pup_Facility_g %>% pull(feature), a = res_a_pup_Facility_g %>% pull(taxa_id)), show.plot = F)
attr(a, "class") = NULL
attr(a, "intersections") = NULL
a

a = gplots::venn(list(m = res_m_pup_TREATMENT_g %>% pull(feature), a = res_a_pup_TREATMENT_g %>% pull(taxa_id)), show.plot = F)
attr(a, "class") = NULL
attr(a, "intersections") = NULL
a

```

#### maaslin2 heatmap genus
```{r include=FALSE}
################## pup treatment
picked_taxa = res_m_pup_TREATMENT_g %>% pull(feature)
taxa_new = taxa %>% mutate(taxa_long = str_c(Kingdom,Phylum,Class,Order,Family,Genus,sep = ";"))

meta_all = phylo_rf@sam_data@.Data %>% as.data.frame()
names(meta_all) = phylo_rf@sam_data@names
meta_all$SampleID = phylo_rf@sam_data@row.names

wmeta = meta_all %>% filter(Generation == "F1") %>% 
  mutate(SamplingTimepoint = factor(SamplingTimepoint, levels = c("Week3","Week4","Week5","Week7","Week9","Week12"), ordered = T)) %>% 
  arrange(SamplingTimepoint,TREATMENT, Facility)
phylo_g = tax_glom(phylo_rf, taxrank = "Genus", NArm = F)
ft = otu_table(phylo_g@otu_table) %>% as.matrix() %>% as.data.frame() 
wft = ft[picked_taxa, wmeta %>% pull(SampleID)]
wft = log10((wft + 1)/colSums(ft[,-1])[1])
wft = wft %>% rownames_to_column(var="id") %>% merge(taxa_new %>% select(id, taxa_long)) %>% select(!id) %>% select(taxa_long, everything())
#### estimate log2FC, looks like Maaslin2 coef is the log2FC, Chlorinated water as ref
coef_and_p = log2FC_lm_estimate(wft, wmeta, taxa_id = "taxa_long",
                       fo = "log_RA ~ TREATMENT + Facility + Sex",
                       Var = "TREATMENT")
#### annotation information 
df_col = wmeta %>% select(SamplingTimepoint, TREATMENT, Facility)
column_ha = HeatmapAnnotation(df = df_col %>% select(SamplingTimepoint, TREATMENT, Facility), 
                              col = list(SamplingTimepoint = col_timepoint_pup,
                                         TREATMENT = col_treatment, 
                                         Facility = col_facility),
                              simple_anno_size = unit(3, "mm"), 
                              annotation_label = c("Time point", "Treatment", "Facility"),
                              annotation_name_gp = gpar(fontsize = 7,fontface = "bold"), 
                              annotation_name_side = "left")
wft2 = wft %>% column_to_rownames(var = "taxa_long")
df_row_name_label = structure(rownames(wft2), names = rownames(wft2))
row_dend = as.dendrogram(hclust(dist(wft2)))
#### the heatmap below don't cluster column(sample), so samples are in orders
p = Heatmap(as.matrix(wft2), 
            col = mat_col_fun, 
            name = "Relative Abundance",
            heatmap_legend_param = list(at = c(-4, -3.3, -2.6, -1.9, -1.2, -0.5), 
                                        labels = c("0.01%", "0.05%", "0.25%", "1.25%", "6.25%", "31.25%"), 
                                        labels_gp = gpar(fontsize = 7), 
                                        title_gp = gpar(fontsize = 8, fontface = "bold")),
            cluster_columns = F, show_column_names = F,
            column_title_gp = gpar(fontsize = 7, fontface = "bold"),
            column_split = df_col$SamplingTimepoint,
            cluster_rows = row_dend,
            row_names_gp = gpar(fontsize = 8),
            row_labels = df_row_name_label,
            row_dend_width = unit(15, "mm"),
            row_names_max_width = max_text_width(df_row_name_label, gp = gpar(fontsize = 8)),
            top_annotation = column_ha,
            height = unit(8, "mm")*dim(wft)[1],
            width = unit(0.5, "mm")*dim(wft)[2]
            )
p1 = draw(p, heatmap_legend_side = "left", annotation_legend_side = "left")
pdf(file = paste0("output/taxa/maaslin2/pup_treatment_by_week_heatmap_g.pdf"), 
    width = ComplexHeatmap:::width(p1) %>% as.numeric()/25.4, 
    height = ComplexHeatmap:::height(p1) %>% as.numeric()/25.4, useDingbats = F)
draw(p1)
dev.off()

row_ord = row_order(p1)
taxa_ord = rownames(wft2)[row_ord]
## all genus coef data
a = lapply(c("Week3","Week4","Week5","Week7","Week9","Week12"),
           function(x){res = read.delim(paste0("output/taxa/maaslin2/pups_",x, "_Genus/all_results.tsv"))
                       res$Group = x
                       res
           })  
res_m_pup_g = do.call(rbind, a) %>% mutate(feature = str_remove(feature,"^X")) %>% 
  left_join(taxa_new %>% select(id, taxa_long),by = c("feature"="id"))

mat1 = res_m_pup_g %>% filter(taxa_long %in% taxa_ord, metadata == "TREATMENT") %>% 
  select(taxa_long, Group, coef) %>% pivot_wider(names_from = "Group", values_from = "coef") %>% 
  column_to_rownames(var = "taxa_long")
mat1 = mat1[taxa_ord,]

mat2 = res_m_pup_g %>% filter(taxa_long %in% taxa_ord, metadata == "TREATMENT") %>% 
  select(taxa_long, Group, qval) %>% 
  mutate(qval = ifelse(qval< 0.1, as.character(round(qval,4)),"")) %>% 
  pivot_wider(names_from = "Group", values_from = "qval") %>% 
  column_to_rownames(var = "taxa_long")
mat2 = mat2[taxa_ord,]

df_row_name_label = structure(rownames(mat1), names = rownames(mat1))
p = Heatmap(as.matrix(mat1), 
            col = col_fun, 
            name = "Log2 Fold Change",
            heatmap_legend_param = list(at = c(-2, -1, 0, 1, 2), 
                                        labels = c("-2", "-1", "0", "1", "2"), 
                                        labels_gp = gpar(fontsize = 7), 
                                        title_gp = gpar(fontsize = 8, fontface = "bold")),
            cluster_columns = F, show_column_names = T,
            column_title_gp = gpar(fontsize = 7, fontface = "bold"),
            cluster_rows = F,
            row_names_gp = gpar(fontsize = 8),
            row_labels = df_row_name_label,
            row_dend_width = unit(15, "mm"),
            row_names_max_width = max_text_width(df_row_name_label, gp = gpar(fontsize = 8)),
            cell_fun = function(j,i, x, y, width, height, fill){
              grid.text(sprintf("%s", mat2[i, j]), x, y, gp = gpar(fontsize = 7))
            },
            height = unit(5, "in"),
            width = unit(10, "in")
            )
p1 = draw(p, heatmap_legend_side = "left", annotation_legend_side = "left")
pdf(file = paste0("output/taxa/maaslin2/pup_treatment_by_week_heatmap_g_qval_fc.pdf"), 
    width = ComplexHeatmap:::width(p1) %>% as.numeric()/25.4, 
    height = ComplexHeatmap:::height(p1) %>% as.numeric()/25.4,
    useDingbats = F)
draw(p1)
dev.off()

############# pup facility
picked_taxa = res_m_pup_Facility_g %>% pull(feature)
picked_taxa = picked_taxa[which(abs(res_m_pup_Facility_g[,-1]) %>% apply(., 1, FUN = min, na.rm = T) > 0.8)]
taxa_new = taxa %>% mutate(taxa_long = str_c(Kingdom,Phylum,Class,Order,Family,Genus,sep = ";"))

meta_all = phylo_rf@sam_data@.Data %>% as.data.frame()
names(meta_all) = phylo_rf@sam_data@names
meta_all$SampleID = phylo_rf@sam_data@row.names

wmeta = meta_all %>% filter(Generation == "F1") %>% 
  mutate(SamplingTimepoint = factor(SamplingTimepoint, levels = c("Week3","Week4","Week5","Week7","Week9","Week12"), ordered = T)) %>% 
  arrange(SamplingTimepoint, Facility,TREATMENT)
phylo_g = tax_glom(phylo_rf, taxrank = "Genus", NArm = F)
ft = otu_table(phylo_g@otu_table) %>% as.matrix() %>% as.data.frame() 
wft = ft[picked_taxa, wmeta %>% pull(SampleID)]
wft = log10((wft + 1)/colSums(ft[,-1])[1])
wft = wft %>% rownames_to_column(var="id") %>% merge(taxa_new %>% select(id, taxa_long)) %>% select(!id) %>% select(taxa_long, everything())

#### annotation information 
df_col = wmeta %>% select(SamplingTimepoint, Facility, TREATMENT)
column_ha = HeatmapAnnotation(df = df_col %>% select(SamplingTimepoint, Facility, TREATMENT), 
                              col = list(SamplingTimepoint = col_timepoint_pup,
                                         TREATMENT = col_treatment, 
                                         Facility = col_facility),
                              simple_anno_size = unit(3, "mm"), 
                              annotation_label = c("Time point", "Facility", "Treatment"),
                              annotation_name_gp = gpar(fontsize = 7,fontface = "bold"), 
                              annotation_name_side = "left")
wft2 = wft %>% column_to_rownames(var = "taxa_long")
df_row_name_label = structure(rownames(wft2), names = rownames(wft2))
row_dend = as.dendrogram(hclust(dist(wft2)))
#### the heatmap below don't cluster column(sample), so samples are in orders
p = Heatmap(as.matrix(wft2), 
            col = mat_col_fun, 
            name = "Relative Abundance",
            heatmap_legend_param = list(at = c(-4, -3.3, -2.6, -1.9, -1.2, -0.5), 
                                        labels = c("0.01%", "0.05%", "0.25%", "1.25%", "6.25%", "31.25%"), 
                                        labels_gp = gpar(fontsize = 7), 
                                        title_gp = gpar(fontsize = 8, fontface = "bold")),
            cluster_columns = F, show_column_names = F,
            column_title_gp = gpar(fontsize = 7, fontface = "bold"),
            column_split = df_col$SamplingTimepoint,
            cluster_rows = row_dend,
            row_names_gp = gpar(fontsize = 8),
            row_labels = df_row_name_label,
            row_dend_width = unit(15, "mm"),
            row_names_max_width = max_text_width(df_row_name_label, gp = gpar(fontsize = 8)),
            top_annotation = column_ha,
            height = unit(8, "mm")*dim(wft)[1],
            width = unit(0.5, "mm")*dim(wft)[2]
            )
p1 = draw(p, heatmap_legend_side = "left", annotation_legend_side = "left")
pdf(file = paste0("output/taxa/maaslin2/pup_facility_by_week_heatmap_g.pdf"), 
    width = ComplexHeatmap:::width(p1) %>% as.numeric()/25.4, 
    height = ComplexHeatmap:::height(p1) %>% as.numeric()/25.4, useDingbats = F)
draw(p1)
dev.off()

row_ord = row_order(p1)
taxa_ord = rownames(wft2)[row_ord]
## all genus coef data
a = lapply(c("Week3","Week4","Week5","Week7","Week9","Week12"),
           function(x){res = read.delim(paste0("output/taxa/maaslin2/pups_",x, "_Genus/all_results.tsv"))
                       res$Group = x
                       res
           })  
res_m_pup_g = do.call(rbind, a) %>% mutate(feature = str_remove(feature,"^X")) %>% 
  left_join(taxa_new %>% select(id, taxa_long),by = c("feature"="id"))

mat1 = res_m_pup_g %>% filter(taxa_long %in% taxa_ord, metadata == "Facility") %>% 
  select(taxa_long, Group, coef) %>% pivot_wider(names_from = "Group", values_from = "coef") %>% 
  column_to_rownames(var = "taxa_long")
mat1 = mat1[taxa_ord,]

mat2 = res_m_pup_g %>% filter(taxa_long %in% taxa_ord, metadata == "Facility") %>% 
  select(taxa_long, Group, qval) %>% 
  mutate(qval = ifelse(qval< 0.1, as.character(round(qval,4)),"")) %>% 
  pivot_wider(names_from = "Group", values_from = "qval") %>% 
  column_to_rownames(var = "taxa_long")
mat2 = mat2[taxa_ord,]

df_row_name_label = structure(rownames(mat1), names = rownames(mat1))
p = Heatmap(as.matrix(mat1), 
            col = col_fun, 
            name = "Log2 Fold Change",
            heatmap_legend_param = list(at = c(-2, -1, 0, 1, 2), 
                                        labels = c("-2", "-1", "0", "1", "2"), 
                                        labels_gp = gpar(fontsize = 7), 
                                        title_gp = gpar(fontsize = 8, fontface = "bold")),
            cluster_columns = F, show_column_names = T,
            column_title_gp = gpar(fontsize = 7, fontface = "bold"),
            cluster_rows = F,
            row_names_gp = gpar(fontsize = 8),
            row_labels = df_row_name_label,
            row_dend_width = unit(15, "mm"),
            row_names_max_width = max_text_width(df_row_name_label, gp = gpar(fontsize = 8)),
            cell_fun = function(j,i, x, y, width, height, fill){
              grid.text(sprintf("%s", mat2[i, j]), x, y, gp = gpar(fontsize = 7))
            },
            height = unit(5, "in"),
            width = unit(10, "in")
            )
p1 = draw(p, heatmap_legend_side = "left", annotation_legend_side = "left")
pdf(file = paste0("output/taxa/maaslin2/pup_facility_by_week_heatmap_g_qval_fc.pdf"), 
    width = ComplexHeatmap:::width(p1) %>% as.numeric()/25.4, 
    height = ComplexHeatmap:::height(p1) %>% as.numeric()/25.4,
    useDingbats = F)
draw(p1)
dev.off()

```

#### maaslin2 heatmap genus by sex
```{r include=FALSE}
taxa_new = taxa %>% mutate(taxa_long = str_c(Kingdom,Phylum,Class,Order,Family,Genus,sep = ";"))
meta_all = phylo_rf@sam_data@.Data %>% as.data.frame()
names(meta_all) = phylo_rf@sam_data@names
meta_all$SampleID = phylo_rf@sam_data@row.names
phylo_g = tax_glom(phylo_rf, taxrank = "Genus", NArm = F)
ft = otu_table(phylo_g@otu_table) %>% as.matrix() %>% as.data.frame() 


a = lapply(paste(rep(c("M","F"),each = 6),rep(c("Week3","Week4","Week5","Week7","Week9","Week12"),2),  sep = "_"),
           function(x){read.csv(paste0("output/taxa/maaslin2/summary_pups_",x, ".csv"))})  
res_all_sig = do.call(rbind, a) 

res_TREATMENT_g = res_all_sig %>% filter(metadata == "TREATMENT",level == "Genus") %>% 
  mutate(Sex = str_split(Group, "_", simplify = T)[,1],
         Timepoint = str_split(Group, "_", simplify = T)[,2]) 

res_Facility_g = res_all_sig %>% filter(metadata == "Facility",level == "Genus") %>% 
  mutate(Sex = str_split(Group, "_", simplify = T)[,1],
         Timepoint = str_split(Group, "_", simplify = T)[,2]) 

a = lapply(paste(rep(c("M","F"),each = 6),rep(c("Week3","Week4","Week5","Week7","Week9","Week12"),2),  sep = "_"),
           function(x){res = read.delim(paste0("output/taxa/maaslin2/pups_",x, "_Genus/all_results.tsv"))
                       res$Group = x
                       res
           })  
res_all_g = do.call(rbind, a) %>% mutate(feature = str_remove(feature,"^X")) %>% 
  left_join(taxa_new %>% select(id, taxa_long),by = c("feature"="id")) %>% 
  mutate(Sex = str_split(Group, "_", simplify = T)[,1],
         Timepoint = str_split(Group, "_", simplify = T)[,2])

for(v in c("TREATMENT", "Facility")){
  for(s in c("M", "F")){
    res_sig = res_all_sig %>% filter(metadata == v,level == "Genus") %>% 
      mutate(Sex = str_split(Group, "_", simplify = T)[,1],
             Timepoint = str_split(Group, "_", simplify = T)[,2])
    if(nrow(res_sig) > 50){
      res_sig$abs_FC = abs(res_sig$coef)
      res_sig = res_sig %>% arrange(desc(abs_FC)) %>% top_n(50) 
      picked_taxa = res_sig %>% filter(Sex == s) %>% pull(feature)
    }else{
      picked_taxa = res_sig %>% filter(Sex == s) %>% pull(feature)
    } 
    wmeta = meta_all %>% filter(Generation == "F1", Sex == s) %>% 
      mutate(SamplingTimepoint = factor(SamplingTimepoint, levels = c("Week3","Week4","Week5","Week7","Week9","Week12"), ordered = T)) 
    if(v == "TREATMENT"){
      wmeta = wmeta %>% arrange(SamplingTimepoint,TREATMENT, Facility)
      df_col = wmeta %>% select(SamplingTimepoint, TREATMENT, Facility)
      column_ha = HeatmapAnnotation(df = df_col, 
                                    col = list(SamplingTimepoint = col_timepoint_pup,
                                                TREATMENT = col_treatment, 
                                                Facility = col_facility),
                                    simple_anno_size = unit(3, "mm"), 
                                    annotation_label = c("Time point", "Treatment", "Facility"),
                                    annotation_name_gp = gpar(fontsize = 7,fontface = "bold"), 
                                    annotation_name_side = "left")
    }else{
      wmeta = wmeta %>% arrange(SamplingTimepoint,Facility,TREATMENT)
      df_col = wmeta %>% select(SamplingTimepoint, Facility,TREATMENT)
      column_ha = HeatmapAnnotation(df = df_col, 
                                    col = list(SamplingTimepoint = col_timepoint_pup,
                                               TREATMENT = col_treatment, 
                                               Facility = col_facility),
                                    simple_anno_size = unit(3, "mm"), 
                                    annotation_label = c("Time point", "Facility","Treatment"),
                                    annotation_name_gp = gpar(fontsize = 7,fontface = "bold"), 
                                    annotation_name_side = "left")
    }
    wft = ft[picked_taxa, wmeta %>% pull(SampleID)]
    wft = log10((wft + 1)/colSums(ft[,-1])[1])
    wft = wft %>% rownames_to_column(var="id") %>% merge(taxa_new %>% select(id, taxa_long)) %>% select(!id) %>% select(taxa_long, everything())
    wft2 = wft %>% column_to_rownames(var = "taxa_long")
    df_row_name_label = structure(rownames(wft2), names = rownames(wft2))
    row_dend = as.dendrogram(hclust(dist(wft2)))
    p = Heatmap(as.matrix(wft2), 
                col = mat_col_fun, 
                name = "Relative Abundance",
                heatmap_legend_param = list(at = c(-4, -3.3, -2.6, -1.9, -1.2, -0.5), 
                                        labels = c("0.01%", "0.05%", "0.25%", "1.25%", "6.25%", "31.25%"), 
                                        labels_gp = gpar(fontsize = 7), 
                                        title_gp = gpar(fontsize = 8, fontface = "bold")),
                cluster_columns = F, show_column_names = F,
                column_title_gp = gpar(fontsize = 7, fontface = "bold"),
                column_split = df_col$SamplingTimepoint,
                cluster_rows = row_dend,
                row_names_gp = gpar(fontsize = 8),
                row_labels = df_row_name_label,
                row_dend_width = unit(15, "mm"),
                row_names_max_width = max_text_width(df_row_name_label, gp = gpar(fontsize = 8)),
                top_annotation = column_ha,
                height = unit(8, "mm")*dim(wft)[1],
                width = unit(0.5, "mm")*dim(wft)[2]
                )
    p1 = draw(p, heatmap_legend_side = "left", annotation_legend_side = "left")
    pdf(file = paste0("output/taxa/maaslin2/pup_", v, "_by_week_Sex_",s ,"_heatmap_g.pdf"), 
        width = ComplexHeatmap:::width(p1) %>% as.numeric()/25.4, 
        height = ComplexHeatmap:::height(p1) %>% as.numeric()/25.4, useDingbats = F)
    draw(p1)
    dev.off()
    row_ord = row_order(p1)
    taxa_ord = rownames(wft2)[row_ord]
    mat1 = res_all_g %>% filter(taxa_long %in% taxa_ord, metadata == v, Sex == s) %>% 
      select(taxa_long, Timepoint, coef) %>% pivot_wider(names_from = "Timepoint", values_from = "coef") %>% 
      column_to_rownames(var = "taxa_long")
    mat1 = mat1[taxa_ord,]
    
    mat2 = res_all_g %>% filter(taxa_long %in% taxa_ord, metadata == v, Sex == s) %>% 
      select(taxa_long, Timepoint, qval) %>% 
      mutate(qval = ifelse(qval< 0.1, as.character(round(qval,4)),"")) %>% 
      pivot_wider(names_from = "Timepoint", values_from = "qval") %>% 
      column_to_rownames(var = "taxa_long")
    mat2 = mat2[taxa_ord,]
    df_row_name_label = structure(rownames(mat1), names = rownames(mat1))
    p = Heatmap(as.matrix(mat1), 
                col = col_fun, 
                name = "Log2 Fold Change",
                heatmap_legend_param = list(at = c(-2, -1, 0, 1, 2), 
                                            labels = c("-2", "-1", "0", "1", "2"), 
                                            labels_gp = gpar(fontsize = 7), 
                                            title_gp = gpar(fontsize = 8, fontface = "bold")),
                cluster_columns = F, show_column_names = T,
                column_title_gp = gpar(fontsize = 7, fontface = "bold"),
                cluster_rows = F,
                row_names_gp = gpar(fontsize = 8),
                row_labels = df_row_name_label,
                row_dend_width = unit(15, "mm"),
                row_names_max_width = max_text_width(df_row_name_label, gp = gpar(fontsize = 8)),
                cell_fun = function(j,i, x, y, width, height, fill){
                  grid.text(sprintf("%s", mat2[i, j]), x, y, gp = gpar(fontsize = 7))
                },
                height = unit(5, "in"),
                width = unit(10, "in")
                )
    p1 = draw(p, heatmap_legend_side = "left", annotation_legend_side = "left")
    pdf(file = paste0("output/taxa/maaslin2/pup_", v, "_by_week_Sex_",s ,"_heatmap_g_qval_fc.pdf"), 
        width = ComplexHeatmap:::width(p1) %>% as.numeric()/25.4, 
        height = ComplexHeatmap:::height(p1) %>% as.numeric()/25.4,
        useDingbats = F)
    draw(p1)
    dev.off()
  }
} 

```


#Taxa associated with facility, separated by water treatment
```{r}
#diwater
meta_p = meta_filtered %>%  filter(TREATMENT == "DIWATER")
output = "/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/"
  for (week in c(meta_p$Week %>% unique())){
    wmeta = meta_p %>% filter(Week == week)
    fit = do_maaslin2(meta = wmeta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                    fixed = c("Facility"), 
                    random = NULL, 
                    ref = c("Facility,Vivarium"),
                    output = paste0(output,"maaslin2_pups_in_DIWATER_", week)
                    )
    res = summarise_maaslin2(
      path = paste0(output,"maaslin2_pups_in_DIWATER_", week), 
      group = paste0("_", week), taxa)
    write.csv(res, file = paste0(output,"maaslin2_summary_pups_in_DIWATER_", week, ".tsv"),quote = F, row.names = F)  
  }

week3_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_DIWATER_Week 3.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week3")
week7_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_DIWATER_Week 7.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week7")
week12_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_DIWATER_Week 12.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week12")

all_genus <- rbind(week3_genus, week7_genus, week12_genus) %>% 
  mutate(Species = case_when(Species %in% c("uncultured_bacterium", "unidentified", 
                                            "uncultured_organism", "s_") ~ "s_",
                             Species == "Dubosiella_newyorkensis" ~ "s_newyorkensis",
                             T ~ paste("s_",Species, sep = "")),
         Genus = case_when(Genus %in% c("uncultured", "g_") ~ "g",
                           T ~ paste("g_",Genus, sep = "")),
         Family = paste("f_", Family, sep = "")) %>% 
  mutate(FGS_string = paste(Family, Genus, Species, sep = "_"))

dat <- all_genus %>% pivot_wider(names_from = Week, values_from = coef) %>% 
  select(ASV, (last_col(3)):last_col()) %>% 
  group_by(ASV, FGS_string) %>%
  summarise(
    Week3  = first(na.omit(Week3)),
    Week7  = first(na.omit(Week7)),
    Week12 = first(na.omit(Week12)),
    .groups = "drop") %>% 
  arrange(desc(Week3), desc(Week7), desc(Week12)) 
dat[12, 2] <- paste("c_Bacilli_", dat[12,2], sep = "")
dat <- column_to_rownames(dat, "FGS_string")
dat <- dat[,-1] #remove ASV col
mat_di <- as.matrix(dat)
dat[is.na(dat)] <- 0 #replace NA with 0

col_fun = circlize::colorRamp2(c(-3, 0, 6), c("blue", "white", "red"))

p <- Heatmap(mat,
             heatmap_legend_param = list(title = "MaAsLin2\nCoefficient",
                                         at = c(-3, 0, 3, 6)),
             row_order = rownames(mat),
             column_order = colnames(mat),
             col = col_fun,
             row_names_side = "left",
             row_names_gp = gpar(fontsize = 8),
             #row_title = "ASV Species",
             column_names_side = "top",
             column_names_rot = 0,
             column_names_centered = TRUE,
             column_names_gp = gpar(fontsize = 10),
             column_labels = c("Week 3", "Week 7", "Week 12"),)

p1 <- draw(p, padding = unit(c(2, 30, 2, 2), "mm"))
pdf(file = "/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/Fig1_PanelD_taxa_252.pdf", width = 10, height = 8, bg = "white")
p1
dev.off()

#write to table for ppt
dat <- all_genus %>% pivot_wider(names_from = Week, values_from = c(coef, N, N.not.0, qval)) %>% 
  select(ASV, (last_col(13)):last_col()) %>% 
  group_by(ASV, FGS_string) %>%
  summarise(
    coef_Week3  = first(na.omit(coef_Week3)),
    N_Week3  = first(na.omit(N_Week3)),
    N.not.0_Week3  = first(na.omit(N.not.0_Week3)),
    qval_Week3  = first(na.omit(qval_Week3)),
    coef_Week7  = first(na.omit(coef_Week7)),
    N_Week7  = first(na.omit(N_Week7)),
    N.not.0_Week7  = first(na.omit(N.not.0_Week7)),
    qval_Week7  = first(na.omit(qval_Week7)),
    coef_Week12  = first(na.omit(coef_Week12)),
    N_Week12  = first(na.omit(N_Week12)),
    N.not.0_Week12  = first(na.omit(N.not.0_Week12)),
    qval_Week12  = first(na.omit(qval_Week12)),
    .groups = "drop")
dat <- dat[,-1] #remove ASV col
write.csv(dat, file = "/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/Fig1_PanelD_taxa_252_table.csv", row.names=F)



#chlorinated
meta_p = meta_filtered %>%  filter(TREATMENT == "CHLORINATED")
output = "/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/"
  for (week in c(meta_p$Week %>% unique())){
    wmeta = meta_p %>% filter(Week == week)
    fit = do_maaslin2(meta = wmeta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                    fixed = c("Facility"), 
                    random = NULL, 
                    ref = c("Facility,Vivarium"),
                    output = paste0(output,"maaslin2_pups_in_CHLORINATED_", week)
                    )
    res = summarise_maaslin2(
      path = paste0(output,"maaslin2_pups_in_CHLORINATED_", week), 
      group = paste0("_", week), taxa)
    write.csv(res, file = paste0(output,"maaslin2_summary_pups_in_CHLORINATED_", week, ".tsv"),quote = F, row.names = F)  
  }

week3_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_CHLORINATED_Week 3.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week3")
week7_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_CHLORINATED_Week 7.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week7")
week12_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_CHLORINATED_Week 12.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week12")

all_genus <- rbind(week3_genus, week7_genus, week12_genus) %>% 
  mutate(Species = case_when(Species %in% c("uncultured_bacterium", "unidentified", 
                                            "uncultured_organism", "s_") ~ "s_",
                             Species == "Dubosiella_newyorkensis" ~ "s_newyorkensis",
                             Species == "Adlercreutzia_mucosicola" ~ "s_mucosicola",
                             T ~ paste("s_",Species, sep = "")),
         Genus = case_when(Genus %in% c("uncultured", "g_") ~ "g",
                           Genus == "Enterorhabdus" ~ "g_Adlercreutzia",
                           T ~ paste("g_",Genus, sep = "")),
         Family = paste("f_", Family, sep = "")) %>% 
  mutate(FGS_string = paste(Family, Genus, Species, sep = "_"))

dat <- all_genus %>% pivot_wider(names_from = Week, values_from = coef) %>% 
  select(ASV, (last_col(3)):last_col()) %>% 
  group_by(ASV, FGS_string) %>%
  summarise(
    Week3  = first(na.omit(Week3)),
    Week7  = first(na.omit(Week7)),
    Week12 = first(na.omit(Week12)),
    .groups = "drop") %>% 
  arrange(desc(Week3), desc(Week7), desc(Week12)) 
dat[10, 2] <- paste("c_Bacilli_", dat[10,2], sep = "")
dat <- column_to_rownames(dat, "FGS_string")
dat <- dat[,-1] #remove ASV col
mat_chlor <- as.matrix(dat)
dat[is.na(dat)] <- 0 #replace NA with 0

col_fun = circlize::colorRamp2(c(-3, 0, 6), c("blue", "white", "red"))

p <- Heatmap(mat,
             heatmap_legend_param = list(title = "MaAsLin2\nCoefficient",
                                         at = c(-3, 0, 3, 6)),
             row_order = rownames(mat),
             column_order = colnames(mat),
             col = col_fun,
             row_names_side = "left",
             row_names_gp = gpar(fontsize = 8),
             #row_title = "ASV Species",
             column_names_side = "top",
             column_names_rot = 0,
             column_names_centered = TRUE,
             column_names_gp = gpar(fontsize = 10),
             column_labels = c("Week 3", "Week 7", "Week 12"))

p1 <- draw(p, padding = unit(c(2, 30, 2, 2), "mm"))
pdf(file = "/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/Fig1_taxa_252_CHLORINATED.pdf", width = 10, height = 8, bg = "white")
p1
dev.off()

#write to table for ppt
dat <- all_genus %>% pivot_wider(names_from = Week, values_from = c(coef, N, N.not.0, qval)) %>% 
  select(ASV, (last_col(13)):last_col()) %>% 
  group_by(ASV, FGS_string) %>%
  summarise(
    coef_Week3  = first(na.omit(coef_Week3)),
    N_Week3  = first(na.omit(N_Week3)),
    N.not.0_Week3  = first(na.omit(N.not.0_Week3)),
    qval_Week3  = first(na.omit(qval_Week3)),
    coef_Week7  = first(na.omit(coef_Week7)),
    N_Week7  = first(na.omit(N_Week7)),
    N.not.0_Week7  = first(na.omit(N.not.0_Week7)),
    qval_Week7  = first(na.omit(qval_Week7)),
    coef_Week12  = first(na.omit(coef_Week12)),
    N_Week12  = first(na.omit(N_Week12)),
    N.not.0_Week12  = first(na.omit(N.not.0_Week12)),
    qval_Week12  = first(na.omit(qval_Week12)),
    .groups = "drop")
dat <- dat[,-1] #remove ASV col
write.csv(dat, file = "/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/Fig1_taxa_252_CHLORINATED_table.csv", row.names=F)




#into a single heatmap
mat_combined <- merge(mat_di, mat_chlor, by = 0, all.x = TRUE, all.y = TRUE)
mat_combined <- column_to_rownames(mat_combined, var = "Row.names") %>% 
  arrange(desc(Week3.x), desc(Week7.x), desc(Week12.x), desc(Week3.y), desc(Week7.y), desc(Week12.y))
mat_combined[is.na(mat_combined)] <- 0

source_anno <- c(
  rep("Diwater", ncol(mat_di)),
  rep("Chlorinated", ncol(mat_chlor))
)
col_anno <- HeatmapAnnotation(
  Treatment = source_anno,
  col = list(
    Treatment = c(
      Diwater = "darkgreen",
      Chlorinated = "#ad61c6"
    )
  )
)
col_fun = circlize::colorRamp2(c(-3, 0, 6), c("blue", "white", "red"))

h <- Heatmap(
  mat_combined,
  width = ncol(mat_combined) * unit(8, "mm"), 
  height = nrow(mat_combined) * unit(4, "mm"),
  heatmap_legend_param = list(title = "MaAsLin2\nCoefficient",
                                         at = c(-3, 0, 3, 6)),
  top_annotation = col_anno,
  row_order = rownames(mat_combined),
   column_order = colnames(mat_combined),
   col = col_fun,
   row_names_side = "left",
   row_names_gp = gpar(fontsize = 8),
   #row_title = "ASV Species",
   column_names_side = "top",
   column_names_rot = 0,
   column_names_centered = TRUE,
   column_names_gp = gpar(fontsize = 10),
   column_labels = c("Week 3", "Week 7", "Week 12", "Week 3", "Week 7", "Week 12"),
)

p1 <- draw(h, padding = unit(c(2, 30, 2, 2), "mm"))
pdf(file = "/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/Fig1_taxa_252_combined.pdf", width = 10, height = 8, bg = "white")
p1
dev.off()
```


#Taxa associated with water treatment, separated by facility
```{r}
#farm
meta_p = meta_filtered %>%  filter(Facility == "Farm")
output = "/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/"
  for (week in c(meta_p$Week %>% unique())){
    wmeta = meta_p %>% filter(Week == week)
    fit = do_maaslin2(meta = wmeta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                    fixed = c("TREATMENT"), 
                    random = NULL, 
                    ref = c("TREATMENT,CHLORINATED"),
                    output = paste0(output,"maaslin2_pups_in_FARM_", week)
                    )
    res = summarise_maaslin2(
      path = paste0(output,"maaslin2_pups_in_FARM_", week), 
      group = paste0("_", week), taxa)
    write.csv(res, file = paste0(output,"maaslin2_summary_pups_in_FARM_", week, ".tsv"),quote = F, row.names = F)  
  }

week3_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_FARM_Week 3.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week3")
week7_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_FARM_Week 7.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week7")
week12_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_FARM_Week 12.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week12")



#vivarium
meta_p = meta_filtered %>%  filter(Facility == "Vivarium")
output = "/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/"
  for (week in c(meta_p$Week %>% unique())){
    wmeta = meta_p %>% filter(Week == week)
    fit = do_maaslin2(meta = wmeta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                    fixed = c("TREATMENT"), 
                    random = NULL, 
                    ref = c("TREATMENT,CHLORINATED"),
                    output = paste0(output,"maaslin2_pups_in_VIVARIUM_", week)
                    )
    res = summarise_maaslin2(
      path = paste0(output,"maaslin2_pups_in_VIVARIUM_", week), 
      group = paste0("_", week), taxa)
    write.csv(res, file = paste0(output,"maaslin2_summary_pups_in_VIVARIUM_", week, ".tsv"),quote = F, row.names = F)  
  }

week3_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_VIVARIUM_Week 3.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week3")
week7_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_VIVARIUM_Week 7.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week7")
week12_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_VIVARIUM_Week 12.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week12")


```



#Maaslin sex-specific taxa, separated by facility and water treatment
```{r}
#diwater
meta_p = meta_filtered %>%  filter(TREATMENT == "DIWATER")
output = "/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/"
  for (week in c(meta_p$Week %>% unique())){
    for(f in c(meta_p$Facility %>% unique())){
      wmeta = meta_p %>% filter(Week == week & Facility == f)
      fit = do_maaslin2(meta = wmeta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                      fixed = c("Sex"), 
                      random = NULL, 
                      ref = c("Sex,Male"),
                      output = paste0(output,"maaslin2_pups_in_DIWATER_", f, "_", week)
                      )
      res = summarise_maaslin2(
        path = paste0(output,"maaslin2_pups_in_DIWATER_", f, "_", week), 
        group = paste0("_", week), taxa)
      write.csv(res, file = paste0(output,"maaslin2_summary_pups_in_DIWATER_", f, "_", week, ".tsv"),quote = F, row.names = F)  
      }
  }

week3_Farm_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_DIWATER_Farm_Week 3.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week3")
week7_Farm_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_DIWATER_Farm_Week 7.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week7")
week12_Farm_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_DIWATER_Farm_Week 12.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week12")
week3_Viv_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_DIWATER_Vivarium_Week 3.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week3")
week7_Viv_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_DIWATER_Vivarium_Week 7.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week7")
week12_Viv_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_DIWATER_Vivarium_Week 12.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week12")

#Turicibater is feature of interest, present in week 7+12 farm
sample_ids <- meta_filtered %>% filter(TREATMENT == "DIWATER" & Week %in% c("Week 7", "Week 12")) %>% 
  mutate(Week = factor(Week,levels = c("Week 3", "Week 7", "Week 12"), ordered = T))
asv_data <- otu_table(phylo_rf) %>% as.data.frame()

feature = week12_Farm_genus[1,1] %>% as.character()
feature_counts <- asv_data[feature,]
feature_counts <- t(feature_counts) %>% as.data.frame() %>% rownames_to_column(var = "SampleID")

feature_filtered <- left_join(sample_ids, feature_counts, by = "SampleID") %>% rename(seq_count = c4a7d3b208ece70a26268344fbb93c52)

g1 <- ggplot(feature_filtered %>% filter(Week == "Week 7" & Facility == "Farm"), aes(x = Sex, y = seq_count, fill = Sex)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  scale_fill_manual(values = col_sex) +
  labs(y = "Sequence Counts") +
       #title = "Week 7, Farm") +
  #facet_grid(Week ~ Facility) +
  theme_bw() + theme_all

g2 <- ggplot(feature_filtered %>% filter(Week == "Week 7" & Facility == "Vivarium"), aes(x = Sex, y = seq_count, fill = Sex)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  scale_fill_manual(values = col_sex) +
  labs(y = "Sequence Counts") +
       #title = "Week 7, Vivarium") +
  #facet_grid(Week ~ Facility) +
  theme_bw() + theme_all

g3 <- ggplot(feature_filtered %>% filter(Week == "Week 12" & Facility == "Farm"), aes(x = Sex, y = seq_count, fill = Sex)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  scale_fill_manual(values = col_sex) +
  labs(y = "Sequence Counts") +
       #title = "Week 12, Farm") +
  #facet_grid(Week ~ Facility) +
  theme_bw() + theme_all

g4 <- ggplot(feature_filtered %>% filter(Week == "Week 12" & Facility == "Vivarium"), aes(x = Sex, y = seq_count, fill = Sex)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  scale_fill_manual(values = col_sex) +
  labs(y = "Sequence Counts") +
       #title = "Week 12, Vivarium") +
  #facet_grid(Week ~ Facility) +
  theme_bw() + theme_all

p <- g2 + g3 +
     g1 + g4 + 
     plot_layout(guides = "collect", ncol = 2) + 
     plot_annotation(title = "", tag_levels = list(c("G", "H", "I", "J"))) & theme(plot.tag = element_text(size = 12, face = "bold"))
ggsave(file = paste0("/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/","turicibacter_diwater_sex_differences_by_age-facility_for_paper.pdf"),
          plot = p, width = 7, height =  5, units = "in", dpi = 300, device = "pdf")




#chlorinated
meta_p = meta_filtered %>%  filter(TREATMENT == "CHLORINATED")
output = "/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/"
  for (week in c(meta_p$Week %>% unique())){
    for(f in c(meta_p$Facility %>% unique())){
      wmeta = meta_p %>% filter(Week == week & Facility == f)
      fit = do_maaslin2(meta = wmeta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                      fixed = c("Sex"), 
                      random = NULL, 
                      ref = c("Sex,Male"),
                      output = paste0(output,"maaslin2_pups_in_CHLORINATED_", f, "_", week)
                      )
      res = summarise_maaslin2(
        path = paste0(output,"maaslin2_pups_in_CHLORINATED_", f, "_", week), 
        group = paste0("_", week), taxa)
      write.csv(res, file = paste0(output,"maaslin2_summary_pups_in_CHLORINATED_", f, "_", week, ".tsv"),quote = F, row.names = F)  
      }
  }

week3_Farm_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_CHLORINATED_Farm_Week 3.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week3")
week7_Farm_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_CHLORINATED_Farm_Week 7.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week7")
week12_Farm_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_CHLORINATED_Farm_Week 12.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week12")
week3_Viv_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_CHLORINATED_Vivarium_Week 3.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week3")
week7_Viv_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_CHLORINATED_Vivarium_Week 7.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week7")
week12_Viv_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_CHLORINATED_Vivarium_Week 12.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week12")

#Turicibater is feature of interest, present in week 7+12 farm
sample_ids <- meta_filtered %>% filter(TREATMENT == "CHLORINATED" & Week %in% c("Week 7", "Week 12")) %>% 
  mutate(Week = factor(Week,levels = c("Week 3", "Week 7", "Week 12"), ordered = T))
asv_data <- otu_table(phylo_rf) %>% as.data.frame()

feature = week12_Farm_genus[3,1] %>% as.character()
feature_counts <- asv_data[feature,]
feature_counts <- t(feature_counts) %>% as.data.frame() %>% rownames_to_column(var = "SampleID")

feature_filtered <- left_join(sample_ids, feature_counts, by = "SampleID") %>% rename(seq_count = c4a7d3b208ece70a26268344fbb93c52)

g1 <- ggplot(feature_filtered %>% filter(Week == "Week 7" & Facility == "Farm"), aes(x = Sex, y = seq_count, fill = Sex)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  scale_fill_manual(values = col_sex) +
  labs(y = "Sequence Counts",
       title = "Week 7, Farm") +
  #facet_grid(Week ~ Facility) +
  theme_bw() + theme_all

g2 <- ggplot(feature_filtered %>% filter(Week == "Week 7" & Facility == "Vivarium"), aes(x = Sex, y = seq_count, fill = Sex)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  scale_fill_manual(values = col_sex) +
  labs(y = "Sequence Counts",
       title = "Week 7, Vivarium") +
  #facet_grid(Week ~ Facility) +
  theme_bw() + theme_all

g3 <- ggplot(feature_filtered %>% filter(Week == "Week 12" & Facility == "Farm"), aes(x = Sex, y = seq_count, fill = Sex)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  scale_fill_manual(values = col_sex) +
  labs(y = "Sequence Counts",
       title = "Week 12, Farm") +
  #facet_grid(Week ~ Facility) +
  theme_bw() + theme_all

g4 <- ggplot(feature_filtered %>% filter(Week == "Week 12" & Facility == "Vivarium"), aes(x = Sex, y = seq_count, fill = Sex)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  scale_fill_manual(values = col_sex) +
  labs(y = "Sequence Counts",
       title = "Week 7, Vivarium") +
  #facet_grid(Week ~ Facility) +
  theme_bw() + theme_all

p <- g1 + g2 +
     g3 + g4 + 
     plot_layout(guides = "collect", ncol = 2) + 
     plot_annotation(title = "", tag_levels = list(c("G", "H", "I", "J"))) & theme(plot.tag = element_text(size = 12, face = "bold"))
ggsave(file = paste0("/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/","turicibacter_diwater_sex_differences_by_age-facility.pdf"),
          plot = p, width = 7, height =  5, units = "in", dpi = 300, device = "pdf")

```


#Maaslin taxa differences between treatments, separated by facility
```{r}
meta_p = meta_filtered
output = "/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/"
  for (week in c(meta_p$Week %>% unique())){
    for(f in c(meta_p$Facility %>% unique())){
      wmeta = meta_p %>% filter(Week == week & Facility == f)
      fit = do_maaslin2(meta = wmeta, sample_id = "SampleID", phylo_rf = phylo_rf, 
                      fixed = c("Treatment"), 
                      random = NULL, 
                      ref = c("Treatment,Chlorinated"),
                      output = paste0(output,"maaslin2_pups_in_", f, "_", week)
                      )
      res = summarise_maaslin2(
        path = paste0(output,"maaslin2_pups_in_", f, "_", week), 
        group = paste0("_", week), taxa)
      write.csv(res, file = paste0(output,"maaslin2_summary_pups_in_", f, "_", week, ".tsv"),quote = F, row.names = F)  
      }
  }

week3_Farm_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_Farm_Week 3.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week3")
week7_Farm_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_Farm_Week 7.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week7")
week12_Farm_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_Farm_Week 12.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week12")
week3_Viv_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_Vivarium_Week 3.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week3")
week7_Viv_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_Vivarium_Week 7.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week7")
week12_Viv_genus <- read_csv('/Users/aiswaryaa/Downloads/CL_rna/20211015_CL_analysis_new/02_taxa_figures/maaslin2_summary_pups_in_Vivarium_Week 12.tsv') %>% filter(level == "Genus") %>% mutate(Week = "Week12")

```
