---
title: "CL_rnaseq"
output: html_document
date: "2026-01-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tximport)
library(DESeq2)
library(rtracklayer)
library(tidyverse)
library(dplyr)
library(rBLAST)
library(ensembldb)
library(EnhancedVolcano)
library(biomaRt)
library(pheatmap)
library(vegan)
library(pathview)
library(gage)
library(gageData)
library(ggvenn)
library(gplots)
library(patchwork)
#library(EnsDb.Mmusculus.v115)
#library(AnnotationHub)
library(clusterProfiler)
library(msigdbr)
library(org.Mm.eg.db)
library(magrittr)
library(DOSE)
library(enrichplot)
library(org.Hs.eg.db)
library(ggnewscale)
library(UpSetR)
library(ComplexUpset)
library(ggrepel)
library(ggpubr)
#library(dendextend)
library(ComplexHeatmap)
library(circlize)
library(readxl)
library(Cairo)

theme_all <- theme(panel.background = element_rect(fill = NA), 
                       plot.title = element_text(hjust = 0.5), 
                       axis.text = element_text(size = 8),
                       axis.title = element_text(size = 8),
                       legend.text = element_text(size=8),
                       legend.title = element_text(size=8, face = "bold"),
                       title = element_text(size=8),
                       aspect.ratio = 1)
```

#Metadata
```{r}
metadata <- read.csv('/Users/aiswaryaa/Downloads/CL_rna/Sampledata_RNAseq_only.txt', header = TRUE, sep = "\t")
metadata$TubeID <- gsub("\\.", "", metadata$TubeID)
metadata <- metadata %>% 
  mutate(Body_Site = str_to_title(Body_Site))

#liver samples
liverMD <- metadata %>% dplyr::filter(Body_Site == "Liver")
liver_samples <- liverMD %>% 
  dplyr::filter(Body_Site=="Liver") %>% 
  .$TubeID

#ileum samples
ileumMD <- metadata %>% dplyr::filter(Body_Site == "Ileum")
ileum_samples <- ileumMD %>% 
  dplyr::filter(Body_Site=="Ileum") %>% 
  .$TubeID

```


#Generate tx2gene mapping file
```{r}
gtf <- rtracklayer::import("/Users/aiswaryaa/Downloads/CL_rna/GRCm39/Mus_musculus.GRCm39.115.gtf")
gtf_df <- as.data.frame(gtf)

tx2gene <- unique(gtf_df[gtf_df$type == "transcript",
                         c("transcript_id", "transcript_version", "gene_id")])

tx2gene$tx <- paste0(tx2gene$transcript_id, ".", tx2gene$transcript_version)

tx2gene <- tx2gene[, c("tx", "gene_id")]
colnames(tx2gene) <- c("tx", "gene")

#verify ID names
#head(tx2gene, 10)

```

#running tximport
```{r}
#liver 
liver.quant.files <- file.path(quant.dir, paste0(liver_samples, "_quant"), "quant.sf")
names(liver.quant.files) <- liver_samples
stopifnot(all(file.exists(liver.quant.files)))

liver_txi <- tximport(liver.quant.files, type = "salmon", tx2gene = tx2gene)
#names(liver_txi)
#head(liver_txi$counts)

#ileum 
ileum.quant.files <- file.path(quant.dir, paste0(ileum_samples, "_quant"), "quant.sf")
names(ileum.quant.files) <- ileum_samples
stopifnot(all(file.exists(ileum.quant.files)))

ileum_txi <- tximport(ileum.quant.files, type = "salmon", tx2gene = tx2gene)
#names(ileum_txi)
#head(ileum_txi$counts)

#save salmon 
#genes, liver and ileum txi
#EXCLUDES OUTLIERS in ileum
save(genelist, tx2gene, liver_txi, ileum_txi, file="/Users/aiswaryaa/Downloads/CL_rna/genes_txis.RData")

```

#Biomart for gene names
```{r}
ensembl <- useEnsembl("genes", dataset="mmusculus_gene_ensembl")

genelist = getBM(attributes = c("ensembl_transcript_id", "ensembl_gene_id", "mgi_symbol", "chromosome_name", "strand", "transcript_start", "transcript_end","entrezgene_id", "entrezgene_accession"), mart = ensembl)

genelist = genelist %>% dplyr::select(ensembl_transcript_id, ensembl_gene_id, mgi_symbol, chromosome_name, strand, transcript_start, transcript_end, entrezgene_id, entrezgene_accession)

x = genelist %>% dplyr::select(ensembl_gene_id, entrezgene_id) %>% unique() %>% dplyr::filter(entrezgene_id!="")

du_ensembl  = x$ensembl_gene_id[duplicated(x$ensembl_gene_id)] %>% unique()
du_entre = x$entrezgene_id[duplicated(x$entrezgene_id)] %>% unique()
x_no_dup = x %>% dplyr::filter(!ensembl_gene_id %in% du_ensembl) %>% dplyr::filter(!entrezgene_id %in% du_entre)
#x_no_dup$ensembl_gene_id %>% unique() %>% length()
#x_no_dup$entrezgene_id %>% unique() %>% length()
y = genelist %>% dplyr::filter(ensembl_gene_id %in% du_ensembl) %>% 
  dplyr::select(ensembl_gene_id, mgi_symbol, chromosome_name, entrezgene_id, entrezgene_accession) %>% unique() %>%
  mutate(check = ifelse(mgi_symbol == entrezgene_accession,"Y","N"))
z = genelist %>% dplyr::filter(entrezgene_id %in% du_entre) %>% 
  dplyr::select(ensembl_gene_id, mgi_symbol, chromosome_name, entrezgene_id, entrezgene_accession) %>% unique() %>%
  mutate(check = ifelse(mgi_symbol == entrezgene_accession,"Y","N"))
yy = y %>% dplyr::filter(check == "Y")
#yy$ensembl_gene_id %>% unique() %>% length()
#yy$entrezgene_id%>% unique() %>% length()
zy = z %>% dplyr::filter(check == "Y")
#zy$ensembl_gene_id %>% unique() %>% length()
#zy$entrezgene_id%>% unique() %>% length()
du_matched = rbind(yy, zy %>% dplyr::filter(!entrezgene_id %in% zy$entrezgene_id[duplicated(zy$entrezgene_id)])) %>% unique()
#du_matched$ensembl_gene_id %>% unique() %>% length()
#du_matched$entrezgene_id %>% unique() %>% length()
x_no_dup_merge_corrected =  rbind(x_no_dup, du_matched %>% dplyr::select(ensembl_gene_id, entrezgene_id))
#x_no_dup_merge_corrected$ensembl_gene_id %>% unique() %>% length()
#x_no_dup_merge_corrected$entrezgene_id %>% unique() %>% length()
genelist = genelist %>% dplyr::select(-entrezgene_id, -entrezgene_accession) %>% left_join(x_no_dup_merge_corrected, by = "ensembl_gene_id")

genelist = genelist %>% mutate(mgi_symbol = ifelse(mgi_symbol=="" | mgi_symbol==" ", ensembl_gene_id, mgi_symbol))

```


#Liver DESeq design and PCA
```{r}
#DESeq design
liverMD$SamplingWeeks <- as.character(liverMD$SamplingWeeks)
#liverMD$Sex <- ifelse(liverMD$Sex=="F", "Female", "Male")
liverMD$Facility <- ifelse(liverMD$Facility=="Farm", "Farm", "Vivarium")
liverMD$age_fac_sex <- paste(liverMD$SamplingWeeks, liverMD$Facility, liverMD$Sex, sep="_")

liver_dds <- DESeqDataSetFromTximport(txi=liver_txi, colData=liverMD, design=~Facility + SamplingWeeks + Sex)
liver_dds <- DESeq(liver_dds)
liver_res <- results(liver_dds)
liver_dds$age_fac_sex <- factor(paste0(liverMD$SamplingWeeks, liverMD$Facility, liverMD$Sex))
design(liver_dds) <- ~ age_fac_sex
liver_dds <- DESeq(liver_dds)
resultsNames(liver_dds)
liver_res <- results(liver_dds)


#PCA
vsdata <- vst(liver_dds, blind=FALSE)
p <- plotPCA(vsdata, intgroup=c("Sex", "SamplingWeeks", "Facility"), returnData=TRUE)
percentVar <- round(100 * attr(p, "percentVar"))
p$age_sex_fac <- paste0("Week ", liverMD$SamplingWeeks, ", ", liverMD$Sex, ", ", liverMD$Facility)
p <- p[order(p$age_sex_fac),]

mat <- t(assay(vsdata))
meta <- as.data.frame(colData(liver_dds))
adonis2(
  mat ~ Sex + SamplingWeeks + Facility, #can also do each variable individually
  data = meta,
  method = "euclidean",
  permutations = 999, 
  by="margin"
)

#Female = salmon, male = blue
#4 weeks = circle, 12 weeks = triangle
#solid = bartlett, hollow = farm
g_pcoa <- ggplot(p, aes(PC1, PC2, color=age_sex_fac)) +
  geom_point(size=3, aes(color=age_sex_fac, shape=age_sex_fac)) +
  scale_shape_manual(values = c(17, 2, 17, 2, 16, 1, 16, 1)) +
  scale_color_manual(values = c("salmon", "salmon", "blue", "blue", "salmon", "salmon", "blue", "blue")) +
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  labs(color = "Group", shape = "Group") +
  theme_bw() + theme_all
  theme(legend.text = element_text(size = 11), axis.title = element_text(size = 12))
  #stat_ellipse() +
  #ggtitle("Liver PCA Plot")
ggsave(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/Liver_PCA_Supp.pdf",
          plot = g, width = 7, height =  5, units = "in", dpi = 300, device = "pdf")
```

#Save liver data
```{r}
#liver metadata, dds, etc. 
save(liverMD, liver_dds, liver_res, file = "/Users/aiswaryaa/Downloads/CL_rna/Liver_data.RData")
```

#Load txi and liver data
```{r}
load("/Users/aiswaryaa/Downloads/CL_rna/genes_txis.RData")
load("/Users/aiswaryaa/Downloads/CL_rna/Liver_data.RData")
```

#Liver volcano plots and heatmaps and MA by sex and age
```{r}
#cat order: age, facility, sex
#function to plot volcano and heatmap
plotting_liver <- function(dds, cat1, cat2) {
  group1 = gsub(" ", "", cat1)
  list1 = strsplit(cat1, split=" ")
  group2 = gsub(" ", "", cat2)
  list2 = strsplit(cat2, split=" ")
  
  note1 = paste0("up in ", list1[[1]][2], " vs ", list2[[1]][2], ", in ", list1[[1]][1], " week old ", list1[[1]][3], " Liver: ")
  note2 = paste0("down in ", list1[[1]][2], " vs ", list2[[1]][2], ", in ", list1[[1]][1], " week old ", list1[[1]][3], " Liver: ")
  
  p_cutoff = 0.05
  fc_cutoff = 1
  
  res <- results(dds, contrast=c("age_fac_sex", group1, group2))
  res_df = as.data.frame(res) %>% mutate(logPvalue = -log10(pvalue))
  
  #Volcano
  up = res_df %>%  dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff) %>% nrow()
  down = res_df %>%  dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff) %>% nrow()
  g = ggplot(res_df %>% dplyr::filter(padj >= p_cutoff), aes(x = log2FoldChange, y = logPvalue)) + 
      geom_point(color = "gray") + 
      geom_point(data = res_df %>% dplyr::filter(padj < p_cutoff), 
                 aes(x = log2FoldChange, y = logPvalue),
                 color = "blue") + 
      geom_point(data = res_df %>% dplyr::filter(padj < p_cutoff, abs(log2FoldChange) > fc_cutoff),
                 aes(x = log2FoldChange, y = logPvalue),
                 color = "red") + 
      geom_point(data = data.frame(log2FoldChange = 0, logPvalue = 0,
                                   category = factor( c("Not significant", "p-value", "p-value and log2 FC"), levels = c("Not significant", "p-value", "p-value and log2 FC"))),
    aes(log2FoldChange, logPvalue, color = category), size = 10, alpha = 0) +
  scale_color_manual(name = "", 
                      values = c("Not significant" = "gray", "p-value" = "blue", "p-value and log2 FC" = "red"), 
                     labels = c("Not significant", "p-value", expression(p-value~and~log[2]~FC))) +
    guides(color = guide_legend(override.aes = list(alpha = 1, size = 4))) +
      labs(x = expression(log[2]~FC), y = expression(-log[10]~(p-value)), 
           title = paste0("Week ", list1[[1]][1], "\nVivarium vs Farm, ", list1[[1]][3], ", Liver")) + 
      theme_bw() + theme(panel.background = element_rect(fill = NA), 
                             plot.title = element_text(hjust = 0.5, size = 16), 
                             text = element_text(color = "black"),
                             axis.text = element_text(size = 14, color = "black"),
                             axis.title = element_text(size = 14),
                             legend.text = element_text(size = 12))
    ymax = ggplot_build(g)$layout$panel_scales_y[[1]]$range$range[2]
    #g = g + geom_text(data = data.frame(log2FoldChange = c(0,0),
    #                             logPvalue = c(ymax*0.95, ymax*0.85),
    #                             note = paste0(c(note1, note2), 
    #                                           c(up, down))),
    #                  aes(x = log2FoldChange, y = logPvalue, label=note), 
    #                  size = 14/.pt, fontface = "bold")
    ggsave(file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/Volcano/", group1 , "_", group2, "_Volcano_Supp.pdf"),
          plot = g, width = 7, height =  5, units = "in", dpi = 300, device = "pdf") 
    
    #MA plot
    g = ggplot(res_df, aes(x = baseMean, y = log2FoldChange)) + 
      geom_point(color = "gray") + 
      geom_point(data = res_df %>% dplyr::filter(padj < p_cutoff), 
                 aes(x = baseMean, y = log2FoldChange),
                 color = "blue") + 
      geom_point(data = res_df %>% dplyr::filter(padj < p_cutoff, abs(log2FoldChange) > fc_cutoff), 
                 aes(x = baseMean, y = log2FoldChange),
                 color = "red") +
      scale_x_log10(breaks = c(1e-1, 1e0, 1e1, 1e2, 1e3, 1e4, 1e5,1e6)) + 
      geom_point(data = data.frame(log2FoldChange = 0, logPvalue = 0,
                                   category = factor( c("Not significant", "Significant (p < 0.05)", "Significant + Large FC\n (p < 0.05 and |LFC| > 1)"), levels = c("Not significant", "Significant (p < 0.05)", "Significant + Large FC\n (p < 0.05 and |LFC| > 1)"))),
    aes(log2FoldChange, logPvalue, color = category), size = 4, alpha = 0) +
  scale_color_manual(name = "Differential Expression", 
                      values = c("Not significant" = "gray", "Significant (p < 0.05)" = "blue", "Significant + Large FC\n (p < 0.05 and |LFC| > 1)" = "red")) +
    guides(color = guide_legend(override.aes = list(alpha = 1, size = 4))) +
      labs(x = "Mean of normalized counts", y = "log2FoldChange (Bartlett/Farm)", title = paste0(list1[[1]][2], " vs ", list2[[1]][2], ", in ", list1[[1]][1], " week old ", list1[[1]][3], " Liver")) + 
      theme_bw() + theme(panel.background = element_rect(fill = NA), 
                             plot.title = element_text(hjust = 0.5), 
                             text = element_text(face = 'bold', color = "black"),
                             axis.text = element_text(size = 14, color = "black"),
                             axis.title = element_text(size = 14))
    ggsave(file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/MA/", group1 , "_", group2, "_MA_plot.pdf"),
               plot = g, width = 7, height =  5, units = "in", dpi = 300, device = "pdf")
    
    
    #Heatmap
    sig <- res[!is.na(res$padj) & res$padj < 0.05 & abs(res$log2FoldChange) > 1, ]
    genes <- rownames(sig)
    length(genes)
    vsd <- vst(dds, blind=FALSE)
    mat <- assay(vsd)[genes, ]
    
    group <- colData(dds)$age_fac_sex
    keep <- group %in% c(group1, group2)
    
    mat <- mat[, keep]
    anno <- data.frame(group=group[keep])
    rownames(anno) <- colnames(mat)
    anno <- mutate(anno, group = ifelse(grepl("Bartlett",anno$group), "Bartlett", "Farm"))
    
    mat_scaled <- t(scale(t(mat))) %>% 
      as.data.frame(.) %>% 
      rownames_to_column(., "gene_id") %>% 
      mutate(., mgi_symbol = genelist$mgi_symbol[match(.$gene_id, genelist$ensembl_gene_id)]) %>% 
      column_to_rownames(., "mgi_symbol") %>% 
      .[, -c(1)]
    
    top <- head(order(sig$padj), 50)
    
    pheatmap(mat_scaled[top, ], annotation_col=anno, main = paste0(list1[[1]][1], " week old ", list1[[1]][3], " Liver"), filename = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/Heatmap/", group1 , "_", group2, "_Heatmap.pdf")) 

    }

plotting_liver(liver_dds,"4 Bartlett Male", "4 Farm Male")
plotting_liver(liver_dds,"4 Bartlett Female", "4 Farm Female") #Heatmap error from no DEG in volcano
plotting_liver(liver_dds,"12 Bartlett Male", "12 Farm Male")
plotting_liver(liver_dds,"12 Bartlett Female", "12 Farm Female")
```

#Liver Volcano plots Combined Sex by age
```{r}
p_cutoff = 0.05
fc_cutoff = 1

#Week 4 with combined sexes
res <- results(liver_dds, contrast = list(c("age_fac_sex_4BartlettMale_vs_12BartlettFemale", "age_fac_sex_4BartlettFemale_vs_12BartlettFemale"), c("age_fac_sex_4FarmMale_vs_12BartlettFemale","age_fac_sex_4FarmFemale_vs_12BartlettFemale")),listValues = c(1/2, -1/2))

res_df <- as.data.frame(res) %>% mutate(logPvalue = -log10(pvalue))%>% 
    rownames_to_column(., "gene_id") 
    res_df <- mutate(res_df, mgi_symbol = genelist$mgi_symbol[match(res_df$gene_id, genelist$ensembl_gene_id)]) %>% 
    mutate(res_df, gene_name = mgi_symbol) %>% 
    drop_na() %>% 
    distinct(mgi_symbol, .keep_all = TRUE) %>% 
    column_to_rownames(., "mgi_symbol") %>% 
    .[, -c(1)]
    
    #genes_of_interest <- c()
    #res_df$label <- if_else(res_df$gene_name %in% genes_of_interest, res_df$gene_name, NA_character_)

  up = res_df %>%  dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff) %>% nrow()
  down = res_df %>%  dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff) %>% nrow()
  g = ggplot(res_df %>% dplyr::filter(padj >= p_cutoff), aes(x = log2FoldChange, y = logPvalue)) + 
      geom_point(color = "gray13") + 
      geom_point(data = res_df %>% dplyr::filter(padj < p_cutoff), 
                 aes(x = log2FoldChange, y = logPvalue),
                 color = "blue") + 
      geom_point(data = res_df %>% dplyr::filter(padj < p_cutoff, abs(log2FoldChange) > fc_cutoff),
                 aes(x = log2FoldChange, y = logPvalue),
                 color = "red") + 
      geom_point(data = data.frame(log2FoldChange = 0, logPvalue = 0,
                                   category = factor( c("Not significant", "p < 0.05", "p < 0.05 and |log2 FC| > 1"), levels = c("Not significant", "p < 0.05", "p < 0.05 and |log2 FC| > 1"))),
    aes(log2FoldChange, logPvalue, color = category), size = 10, alpha = 0) +
  scale_color_manual(name = "", 
                      values = c("Not significant" = "gray13", "p < 0.05" = "blue", "p < 0.05 and |log2 FC| > 1" = "red"), 
                     labels = c("Not significant", "p < 0.05", "p < 0.05 and")) +
    guides(color = guide_legend(override.aes = list(alpha = 1, size = 4))) +
      labs(x = expression(log[2]~FC(Vivarium/Farm)), y = expression(-log[10]~(p-value)), 
           title = paste0("Week 4, Liver")) + 
      theme(plot.title = element_text(hjust = 0.5, size = 16), 
                             text = element_text(color = "black"),
                             axis.text = element_text(size = 16, color = "black"),
                             axis.title = element_text(size = 16),
                             legend.text = element_text(size = 16))
    ymax = ggplot_build(g)$layout$panel_scales_y[[1]]$range$range[2]
g

ggsave(file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/RUMP_Week4_Liver_Volcano.pdf"), plot = g, width = 7, height =  5, units = "in", dpi = 300, device = "pdf")


    
#Week 12 with combined sexes
    res <- results(liver_dds, contrast = list(c("age_fac_sex_12BartlettMale_vs_12BartlettFemale"), c("age_fac_sex_12FarmMale_vs_12BartlettFemale","age_fac_sex_12FarmFemale_vs_12BartlettFemale")),listValues = c(1/2, -1/2))
    
    res_df <- as.data.frame(res) %>% mutate(logPvalue = -log10(pvalue))%>% 
    rownames_to_column(., "gene_id") 
    res_df <- mutate(res_df, mgi_symbol = genelist$mgi_symbol[match(res_df$gene_id, genelist$ensembl_gene_id)]) %>% 
    mutate(res_df, gene_name = mgi_symbol) %>% 
    drop_na() %>% 
    distinct(mgi_symbol, .keep_all = TRUE) %>% 
    column_to_rownames(., "mgi_symbol") %>% 
    .[, -c(1)]
    
    genes_of_interest <- c("Cyp2a5", "Cyp46a1", "Cyp3a16", "Cyp2g1")
    res_df$label <- if_else(res_df$gene_name %in% genes_of_interest, res_df$gene_name, NA_character_)

  up = res_df %>%  dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff) %>% nrow()
  down = res_df %>%  dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff) %>% nrow()
  g = ggplot(res_df %>% dplyr::filter(padj >= p_cutoff), aes(x = log2FoldChange, y = logPvalue)) + 
      geom_point(color = "gray13") + 
      geom_point(data = res_df %>% dplyr::filter(padj < p_cutoff), 
                 aes(x = log2FoldChange, y = logPvalue),
                 color = "blue") + 
      geom_point(data = res_df %>% dplyr::filter(padj < p_cutoff, abs(log2FoldChange) > fc_cutoff),
                 aes(x = log2FoldChange, y = logPvalue),
                 color = "red") + 
      geom_point(data = data.frame(log2FoldChange = 0, logPvalue = 0,
                                   category = factor( c("Not significant", "p < 0.05", "p < 0.05 and |log2 FC| > 1"), levels = c("Not significant", "p < 0.05", "p < 0.05 and |log2 FC| > 1"))),
    aes(log2FoldChange, logPvalue, color = category), size = 10, alpha = 0) +
  scale_color_manual(name = "", 
                      values = c("Not significant" = "gray13", "p < 0.05" = "blue", "p < 0.05 and |log2 FC| > 1" = "red"), 
                     labels = c("Not significant", "p < 0.05", "p < 0.05 and")) +
    guides(color = guide_legend(override.aes = list(alpha = 1, size = 4))) +
      labs(x = expression(log[2]~FC(Vivarium/Farm)), y = expression(-log[10]~(p-value)), 
           title = paste0("Week 12, Liver")) + 
      theme(plot.title = element_text(hjust = 0.5, size = 16), 
                             text = element_text(color = "black"),
                             axis.text = element_text(size = 16, color = "black"),
                             axis.title = element_text(size = 16),
                             legend.text = element_text(size = 16)) +
    geom_label_repel(
    data = res_df %>% dplyr::filter(!is.na(label)),
    aes(x = log2FoldChange, y = logPvalue, label = label),
    size = 5,
    show.legend = FALSE,
    min.segment.length = 0
  )
    ymax = ggplot_build(g)$layout$panel_scales_y[[1]]$range$range[2]
g

ggsave(file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/RUMP_Week12_Liver_Volcano.pdf"), plot = g, width = 7, height =  5, units = "in", dpi = 300, device = "pdf")

up_sig_res <- res_df %>% dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff)
named <- as.data.frame(up_sig_res) %>% mutate(logPvalue = -log10(pvalue))%>% 
    rownames_to_column(., "gene_id") 
named <- mutate(named, mgi_symbol = genelist$mgi_symbol[match(named$gene_id, genelist$ensembl_gene_id)]) %>% 
    drop_na() %>% 
    distinct(mgi_symbol, .keep_all = TRUE) %>% 
    column_to_rownames(., "mgi_symbol") %>% 
    .[, -c(1)]
  
down_sig_res <- res_df %>% dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff)
named <- as.data.frame(down_sig_res) %>% mutate(logPvalue = -log10(pvalue))%>% 
    rownames_to_column(., "gene_id") 
  named <- mutate(named, mgi_symbol = genelist$mgi_symbol[match(named$gene_id, genelist$ensembl_gene_id)]) %>% 
    drop_na() %>% 
    distinct(mgi_symbol, .keep_all = TRUE) %>% 
    column_to_rownames(., "mgi_symbol") %>% 
    .[, -c(1)]
```

#Liver Heatmap Main Figure
```{r}

res <- results(liver_dds, contrast=c("age_fac_sex", "4BartlettMale", "4FarmMale"))
male4_sig <- res[!is.na(res$padj) & res$padj < 0.05 & abs(res$log2FoldChange) > 2, ] %>% as.data.frame() %>% arrange(padj) %>% slice(1:30) %>% rownames()

res <- results(liver_dds, contrast=c("age_fac_sex", "12BartlettMale", "12FarmMale"))
male12_sig <- res[!is.na(res$padj) & res$padj < 0.05 & abs(res$log2FoldChange) > 2, ] %>% as.data.frame() %>% arrange(padj) %>% slice(1:30) %>% rownames()

res <- results(liver_dds, contrast=c("age_fac_sex", "4BartlettFemale", "4FarmFemale"))
female4_sig <- res[!is.na(res$padj) & res$padj < 0.05 & abs(res$log2FoldChange) > 2, ] %>% as.data.frame() %>% arrange(padj) %>% slice(1:30) %>% rownames()

res <- results(liver_dds, contrast=c("age_fac_sex", "12BartlettFemale", "12FarmFemale"))
female12_sig <- res[!is.na(res$padj) & res$padj < 0.05 & abs(res$log2FoldChange) > 2, ] %>% as.data.frame() %>% arrange(padj) %>% slice(1:30) %>% rownames()

genes <- c(male4_sig, male12_sig, female4_sig, female12_sig)
#lengths(genes) = 64

vsd <- vst(liver_dds, blind=FALSE)
mat <- assay(vsd)[genes, ]

group <- colData(liver_dds)$age_fac_sex
keep <- group %in% c("4BartlettMale", "4FarmMale", "12BartlettMale", "12FarmMale", "4BartlettFemale", "4FarmFemale", "12BartlettFemale", "12FarmFemale")

mat <- mat[, keep]
anno <- data.frame(group=group[keep])
rownames(anno) <- colnames(mat)
anno <- mutate(anno, Sex = ifelse(grepl("Female",anno$group), "Female", "Male"),
               Age = ifelse(grepl("4",anno$group), "4 weeks", "12 weeks"),
               Facility = ifelse(grepl("Bartlett",anno$group), "Vivarium", "Farm")) %>% 
  .[, -1]

ann_colors = list(
  Age = c("4 weeks" = "gold", "12 weeks" = "darkorange"),
  Sex = c(Female = "firebrick2", Male = "darkmagenta"),
  Facility = c(Farm="forestgreen", Vivarium="dodgerblue"))

#New original heatmap
liverMD$Facility <- ifelse(liverMD$Facility=="Farm", "Farm", "Vivarium")
temp_meta <- liverMD[order(liverMD$Facility), ]
temp_meta <- mutate(temp_meta, SamplingWeeks = ifelse(grepl("4",temp_meta$SamplingWeeks), "4 weeks", "12 weeks")) 
mat <- mat[, temp_meta$TubeID]

col_fun = colorRamp2(c(-4, 0, 4), c("royalblue", "lemonchiffon", "red1"))
#highlights cells with p < .1
#col_fun = colorRamp2(c(-4, -1.3, 0, 1.3, 4), c("blue", "white", "white", "white", "red"))

col_anno <- HeatmapAnnotation(
  Age = temp_meta$SamplingWeeks,
  Sex = temp_meta$Sex,
  Facility = temp_meta$Facility,
  annotation_name_gp = gpar(fontsize = 10),
  col = list(
    Facility = c("Vivarium" = "dodgerblue", "Farm" = "forestgreen"),
    Age = c("4 weeks" = "gold", "12 weeks" = "darkorange"),
    Sex = c("Male" = "darkmagenta", "Female" = "firebrick2")
  )
)

mat_scaled <- t(scale(t(mat))) %>% 
  as.data.frame(.) %>% 
  rownames_to_column(., "gene_id") %>% 
  mutate(., mgi_symbol = genelist$mgi_symbol[match(.$gene_id, genelist$ensembl_gene_id)]) %>% 
  drop_na() %>% 
  column_to_rownames(., "mgi_symbol") %>% 
  .[, -c(1)] %>% 
  as.matrix()

p <- Heatmap(
  mat_scaled,
  #col = col_fun,
  name = "Z-score",
  top_annotation = col_anno,
  cluster_columns = TRUE,      
  show_column_dend = TRUE,      
  show_row_dend = TRUE,
  column_title = NULL,
  show_column_names = FALSE,
  row_names_gp = gpar(fontsize = 8),
  column_title_side = "bottom"
)

pdf(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/Liver_Heatmap_main2.pdf", width = 7, height = 8)
p
dev.off()

#Supervised by facility
col_anno <- HeatmapAnnotation(
  Facility = temp_meta$Facility,
  Age = temp_meta$SamplingWeeks,
  Sex = temp_meta$Sex,
  annotation_name_gp = gpar(fontsize = 10),
  col = list(
    Facility = c("Vivarium" = "#252525", "Farm" = "#1b9e77"),
    Age = c("4 weeks" = "#f9966d", "12 weeks" = "navy"),
    Sex = c("Male" = "#2b8cbe", "Female" = "#f1726d")
  )
)

p <- Heatmap(
  mat_scaled,
  width = ncol(mat_scaled) * unit(2, "mm"), 
  height = nrow(mat_scaled) * unit(3, "mm"),
  name = "Z-score",
  top_annotation = col_anno,
  cluster_columns = TRUE,       # unsupervised clustering within each block
  column_split = temp_meta$Facility,        # separate blocks by facility
  show_column_dend = TRUE,      # dendrograms for each block
  show_row_dend = TRUE,
  column_title = "Differentially Expressed Genes in Liver",
  show_column_names = FALSE,
  row_names_gp = gpar(fontsize = 8),
  column_title_side = "top",
  column_title_gp = gpar(fontface = "bold")
)

cairo_pdf(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/Liver_Heatmap_facility_supervised.pdf", width = 8, height = 10, fallback_resolution = 300)
p
dev.off()


#Supervised by sex
col_anno <- HeatmapAnnotation(
  Sex = temp_meta$Sex,
  Age = temp_meta$SamplingWeeks,
  Facility = temp_meta$Facility,
  annotation_name_gp = gpar(fontsize = 10),
  col = list(
    Sex = c("Male" = "darkmagenta", "Female" = "firebrick2"),
    Facility = c("Vivarium" = "dodgerblue", "Farm" = "forestgreen"),
    Age = c("4 weeks" = "gold", "12 weeks" = "darkorange")
  )
)

p <- Heatmap(
  mat_scaled,
  name = "Z-score",
  top_annotation = col_anno,
  cluster_columns = TRUE,       # unsupervised clustering within each block
  column_split = temp_meta$Sex,        # separate blocks by facility
  show_column_dend = TRUE,      # dendrograms for each block
  show_row_dend = TRUE,
  column_title = NULL,
  show_column_names = FALSE,
  row_names_gp = gpar(fontsize = 8),
  column_title_side = "bottom"
)

pdf(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/Liver_Heatmap_sex_supervised.pdf", width = 8, height = 6)
p
dev.off()
```


#Liver UpSet plot and venn diagram 
```{r}

p_cutoff = 0.05
fc_cutoff = 1
age_fac_color = c("#7fc97f","#beaed4","#fdc086", "#ffff99")

#up in Bartlett
res <- results(liver_dds, contrast=c("age_fac_sex", "4BartlettMale", "4FarmMale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff)
UP_4BM <- rownames(res_df)

res <- results(liver_dds, contrast=c("age_fac_sex", "12BartlettMale", "12FarmMale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff)
UP_12BM <- rownames(res_df)

res <- results(liver_dds, contrast=c("age_fac_sex", "12BartlettFemale", "12FarmFemale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff)
UP_12BF <- rownames(res_df)

gene_sets <- list(
  "Week 4 Male" = UP_4BM,
  "Week 12 Male" = UP_12BM,
  "Week 4 Female" = c(),
  "Week 12 Female" = UP_12BF
  )

#run this in R console for it to work
plotObject <- UpSetR::upset(fromList(gene_sets), text.scale = c(1,1,1,1,2,2))
pdf(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/Upset_Venn_diagram/Upset_up_in_Bartlett.pdf", width = 8, height = 6)
plotObject
dev.off()

#Venn diagram
g <- ggvenn(gene_sets, fill_color = age_fac_color, set_name_size = 5, text_size = 5, show_percentage = F) + 
  labs(title="Genes Upregulated in Vivarium Mice")
g <- g + theme(plot.title = element_text(hjust = 0.5, size = 14))
ggsave(file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/Upset_Venn_diagram/Venn_up_in_Bartlett_supp.pdf"),
         plot = g, width = 7, height =  5, units = "in", dpi = 300, device = "pdf")




#down in Bartlett
res <- results(liver_dds, contrast=c("age_fac_sex", "4BartlettMale", "4FarmMale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff)
Down_4BM <- rownames(res_df)

res <- results(liver_dds, contrast=c("age_fac_sex", "12BartlettMale", "12FarmMale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff)
Down_12BM <- rownames(res_df)

res <- results(liver_dds, contrast=c("age_fac_sex", "12BartlettFemale", "12FarmFemale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff)
Down_12BF <- rownames(res_df)

gene_sets <- list(
  "Week 4 Male" = Down_4BM,
  "Week 12 Male" = Down_12BM,
  "Week 4 Female" = c(),
  "Week 12 Female" = Down_12BF
  )

#run this in R console for it to work
plotObject <- UpSetR::upset(fromList(gene_sets), text.scale = c(1,1,1,1,2,2))
pdf(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/Upset_Venn_diagram/Upset_down_in_Bartlett.pdf", width = 8, height = 6)
plotObject
dev.off()

#Venn diagram
g <- ggvenn(gene_sets, fill_color = age_fac_color, set_name_size = 5, text_size = 5, show_percentage = F) +
  labs(title="Genes Downregulated in Vivarium Mice")
g <- g + theme(plot.title = element_text(hjust = 0.5, size =14))
ggsave(file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/Upset_Venn_diagram/Venn_down_in_Bartlett_supp.pdf"),
         plot = g, width = 7, height =  5, units = "in", dpi = 300, device = "pdf")

```


#Liver GAGE/pathview pathways
```{r}
data(kegg.sets.mm)
data(sigmet.idx.mm)
kegg.sets.mm = kegg.sets.mm[sigmet.idx.mm]

liver_pathways <- function(dds, cat1, cat2){
    group1 = gsub(" ", "", cat1)
    list1 = strsplit(cat1, split=" ")
    group2 = gsub(" ", "", cat2)
    list2 = strsplit(cat2, split=" ")
    pathway = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/pathways/", group1, "_vs_", group2)
    setwd(pathway)
    
    res_df <- results(dds, contrast=c("age_fac_sex", group1, group2)) %>% 
    as.data.frame() %>% rownames_to_column(var = "ensembl_gene_id") %>% 
    left_join(genelist %>% dplyr::select(ensembl_gene_id, entrezgene_id) %>% unique(), by = "ensembl_gene_id") %>% 
    dplyr::filter(!is.na(entrezgene_id))
    deseq2.fc = res_df$log2FoldChange
    names(deseq2.fc) = res_df$entrezgene_id
    fc.kegg.p = gage(deseq2.fc, gsets = kegg.sets.mm, ref =NULL, samp = NULL)
    
    #New code
    greater_res = fc.kegg.p$greater %>%
      as.data.frame() %>%
      dplyr::filter(!is.na(q.val) & q.val < 0.1) %>%
      dplyr::mutate(direction = "Upregulated") %>%
      dplyr::select(q.val, set.size, stat.mean, direction)
    
    less_res = fc.kegg.p$less %>%
      as.data.frame() %>%
      dplyr::filter(!is.na(q.val) & q.val < 0.1) %>%
      dplyr::mutate(direction = "Downregulatex") %>%
      dplyr::select(q.val, set.size, stat.mean, direction)
    
    res = dplyr::bind_rows(greater_res, less_res)
    res$path_id = str_sub(rownames(res),start = 1, end = 8)
    res = res %>% rownames_to_column(var = "path")
    out.suffix = paste0("Vivarium vs Farm in Week 12 Male Liver")
    setwd("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/pathways/Supp_12BartlettMale_vs_12FarmMale")
    write.csv(res, file = paste0(out.suffix, ".csv"))
    if(nrow(res)>0){
        pv.out.list = sapply(res$path_id, 
                             function(pid) {
                               pathview(gene.data = deseq2.fc, pathway.id = pid, kegg.native = T, 
                                        same.layer = F, species = "mmu",out.suffix=out.suffix
                                        )})}
}  

liver_pathways(dds, "4 Bartlett Male", "4 Farm Male")
liver_pathways(dds,"4 Bartlett Female", "4 Farm Female")
liver_pathways(dds,"12 Bartlett Male", "12 Farm Male")
liver_pathways(dds,"12 Bartlett Female", "12 Farm Female")

getwd()
```


#Liver preparation for GSEA
```{r}
#msigdbr_species()

vsd <- vst(liver_dds, blind=FALSE)
b= "Liver"
for (grp in c("4_Bartlett_Female", "4_Bartlett_Male", "12_Bartlett_Female", "12_Bartlett_Male", "4_Farm_Female", "4_Farm_Male", "12_Farm_Female", "12_Farm_Male" )){
    mat = assay(vsd) %>% as.matrix()
    #colnames(mat) = colData(vsd)$TubeID
    mat = as.data.frame(mat) %>% rownames_to_column(var = "ensembl_gene_id") %>% 
      left_join(genelist %>% dplyr::select(ensembl_gene_id, entrezgene_id) %>% unique() , 
                by = "ensembl_gene_id") %>% 
      dplyr::filter(!is.na(entrezgene_id)) %>%
      dplyr::select(-ensembl_gene_id) %>% column_to_rownames(var = "entrezgene_id") %>% as.matrix()    
    meta_c = liverMD[str_detect(liverMD$age_fac_sex, grp),]
    mat = mat[,meta_c$TubeID]
    out <- mat %>% as.data.frame() %>% 
      rownames_to_column(var = "NAME") %>% 
      mutate(Description = "") %>% 
      dplyr::select(NAME, Description, everything())
    
    con <- file(paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/for_gsea/", b, "_", grp, "_for_gsea.txt"), open="wt")
    #writeLines(paste0("# test"), con)
    write.table(out, con, sep = "\t", quote = F, row.names = F)   
    close(con)
    
    meta_c <- meta_c %>% mutate(code = ifelse(Facility == "Bartlett", 0, 1))
    con <- file(paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/for_gsea/", b, "_", grp, "_for_gsea.cls"), open="wt")
    writeLines(paste(nrow(meta_c),  meta_c$code %>% unique() %>% length(), 1,sep = "\t"), con)
    writeLines(paste("#", "Bartlett", "Farm",sep = "\t"), con)
    writeLines(paste0(as.character(meta_c$code), collapse = "\t"), con)
    close(con)
  }

```

#Liver GSEA, rank by LFC
```{r}

b = "Liver"
res_kegg <- list()
res_go <- list()
res_gsea <- list()

#sex_time_color = c("#7fc97f","#beaed4","#fdc086", "#ffff99")

#HARDCODE EACH GROUP  
cat1="12 Bartlett Female"
cat2="12 Farm Female"
  group1 = gsub(" ", "", cat1)
  list1 = strsplit(cat1, split=" ") #age fac sex
  group2 = gsub(" ", "", cat2)
  list2 = strsplit(cat2, split=" ")
  grp = paste0(list1[[1]][1], "_", list1[[1]][3])

  
  res_df <-  results(liver_dds, contrast = c("age_fac_sex", group1, group2)) %>% 
      as.data.frame() %>% rownames_to_column(var = "ensembl_gene_id") %>% 
      left_join(genelist %>% dplyr::select(ensembl_gene_id, entrezgene_id) %>% unique(), 
                by = "ensembl_gene_id") %>% 
      dplyr::filter(!is.na(entrezgene_id), !is.na(log2FoldChange))
    
    d <- res_df %>% dplyr::select(entrezgene_id, log2FoldChange) %>% 
      arrange(desc(log2FoldChange))
    geneList <- d[,2]
    names(geneList) <- as.character(d[,1])
    
    gsea_go <- gseGO(geneList = geneList,
                     OrgDb = org.Mm.eg.db,
                     ont = "ALL",
                     keyType = "ENTREZID",
                     pAdjustMethod = "BH",
                     pvalueCutoff  = 0.05,
                     eps = 0)
    res <- gsea_go@result %>% mutate(Tissue = b, grp = grp)
    res_go <- rbind(res_go, res)
    res_gsea[["go"]][[b]][[grp]] <- gsea_go
    
    gsea_kegg <- gseKEGG(geneList = geneList,
                         organism = "mmu",
                         keyType = "ncbi-geneid",
                         pAdjustMethod = "BH",
                         pvalueCutoff  = 0.05,
                         eps = 0)
    
    res <- gsea_kegg@result %>% mutate(Tissue = b, grp = grp)
    res_kegg <- rbind(res_kegg, res)
    res_gsea[["kegg"]][[b]][[grp]] <-  gsea_kegg    


#liver_gsea(liver_dds, "4 Bartlett Male", "4 Farm Male")
#liver_gsea(liver_dds, "12 Bartlett Male", "12 Farm Male")
#liver_gsea(liver_dds, "4 Bartlett Female", "4 Farm Female")
#liver_gsea(liver_dds, "12 Bartlett Female", "12 Farm Female")

res_go_all <- dplyr::bind_rows(res_go) %>% mutate(Tissue = "Liver_Bartlett_vs_Farm")
write.table(res_go_all, file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/clusterProfiler/gsea_go_all_LFC.tsv"),
          quote = F, row.names = F,sep = "\t")
as_table <- res_go_all[, c(2,3,5,6,8,13,14)]
write.table(as_table, file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/clusterProfiler/as_table_gsea_go_all_LFC.tsv"),
          quote = F, row.names = F,sep = "\t")

res_kegg_all <- dplyr::bind_rows(res_kegg) %>% mutate(Tissue = "Liver_Bartlett_vs_Farm")
write.table(res_kegg_all, file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/clusterProfiler/gsea_kegg_all_LFC.tsv"),
          quote = F, row.names = F,sep = "\t")
as_table <- res_kegg_all[, c(1,2,4,5,7,12,13)]
write.table(as_table, file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/clusterProfiler/as_table_gsea_kegg_all_LFC.tsv"),
          quote = F, row.names = F,sep = "\t")

save(res_gsea, res_go_all, res_kegg_all, file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/gsea.Rdata")

```

#Liver GSEA, rank by LFC*(-log(p-value))
```{r}

b = "Liver"
res_kegg <- list()
res_go <- list()
res_gsea <- list()

#sex_time_color = c("#7fc97f","#beaed4","#fdc086", "#ffff99")

#HARDCODE EACH GROUP  
cat1="12 Bartlett Female"
cat2="12 Farm Female"
  group1 = gsub(" ", "", cat1)
  list1 = strsplit(cat1, split=" ") #age fac sex
  group2 = gsub(" ", "", cat2)
  list2 = strsplit(cat2, split=" ")
  grp = paste0(list1[[1]][1], "_", list1[[1]][3])

  
  res_df <-  results(liver_dds, contrast = c("age_fac_sex", group1, group2)) %>% 
      as.data.frame() %>% rownames_to_column(var = "ensembl_gene_id") %>% 
      left_join(genelist %>% dplyr::select(ensembl_gene_id, entrezgene_id) %>% unique(), 
                by = "ensembl_gene_id") %>% 
      dplyr::filter(!is.na(entrezgene_id), !is.na(log2FoldChange), !is.na(pvalue))
    
    d <- res_df %>% dplyr::select(entrezgene_id, log2FoldChange) %>% mutate(rank_val = (res_df$log2FoldChange * -log(res_df$pvalue))) %>% 
      arrange(desc(rank_val))
    geneList <- d[,3]
    names(geneList) <- as.character(d[,1])
    
    gsea_go <- gseGO(geneList = geneList,
                     OrgDb = org.Mm.eg.db,
                     ont = "ALL",
                     keyType = "ENTREZID",
                     pAdjustMethod = "BH",
                     pvalueCutoff  = 0.05,
                     eps = 0)
    res <- gsea_go@result %>% mutate(Tissue = b, grp = grp)
    res_go <- rbind(res_go, res)
    res_gsea[["go"]][[b]][[grp]] <- gsea_go
    
    gsea_kegg <- gseKEGG(geneList = geneList,
                         organism = "mmu",
                         keyType = "ncbi-geneid",
                         pAdjustMethod = "BH",
                         pvalueCutoff  = 0.05,
                         eps = 0)
    
    res <- gsea_kegg@result %>% mutate(Tissue = b, grp = grp)
    res_kegg <- rbind(res_kegg, res)
    res_gsea[["kegg"]][[b]][[grp]] <-  gsea_kegg    


#liver_gsea(liver_dds, "4 Bartlett Male", "4 Farm Male")
#liver_gsea(liver_dds, "12 Bartlett Male", "12 Farm Male")
#liver_gsea(liver_dds, "4 Bartlett Female", "4 Farm Female")
#liver_gsea(liver_dds, "12 Bartlett Female", "12 Farm Female")

res_go_all <- dplyr::bind_rows(res_go) %>% mutate(Tissue = "Liver_Bartlett_vs_Farm")
write.table(res_go_all, file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/clusterProfiler/gsea_go_all_LFCxlogPV.tsv"),
          quote = F, row.names = F,sep = "\t")
as_table <- res_go_all[, c(2,3,5,6,8,13,14)]
write.table(as_table, file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/clusterProfiler/as_table_gsea_go_all_LFCxlogPV.tsv"),
          quote = F, row.names = F,sep = "\t")

res_kegg_all <- dplyr::bind_rows(res_kegg) %>% mutate(Tissue = "Liver_Bartlett_vs_Farm")
write.table(res_kegg_all, file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/clusterProfiler/gsea_kegg_all_LFCxlogPV.tsv"),
          quote = F, row.names = F,sep = "\t")
as_table <- res_kegg_all[, c(1,2,4,5,7,12,13)]
write.table(as_table, file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/clusterProfiler/as_table_gsea_kegg_all_LFCxlogPV.tsv"),
          quote = F, row.names = F,sep = "\t")

```


#Liver Supp figure
#PCA, volcano plots(patchwork), venn diagrams
```{r}

liverMD$SamplingWeeks <- as.character(liverMD$SamplingWeeks)
#liverMD$Sex <- ifelse(liverMD$Sex=="F", "Female", "Male")
liverMD$Facility <- ifelse(liverMD$Facility=="Farm", "Farm", "Vivarium")
liverMD$age_fac_sex <- paste(liverMD$SamplingWeeks, liverMD$Facility, liverMD$Sex, sep="_")

#PCA
vsdata <- vst(liver_dds, blind=FALSE)
p <- plotPCA(vsdata, intgroup=c("Sex", "SamplingWeeks", "Facility"), returnData=TRUE)
percentVar <- round(100 * attr(p, "percentVar"), 2)
p$age_sex_fac <- paste0("Week ", liverMD$SamplingWeeks, ", ", liverMD$Sex, ", ", liverMD$Facility)
p <- p[order(p$age_sex_fac),]

mat <- t(assay(vsdata))
meta <- as.data.frame(colData(liver_dds))

#Female = salmon, male = blue
#4 weeks = circle, 12 weeks = triangle
#solid = bartlett, hollow = farm
g_pcoa <- ggplot(p, aes(PC1, PC2, color=age_sex_fac)) +
  geom_point(size=3, aes(color=age_sex_fac, shape=age_sex_fac)) +
  scale_shape_manual(values = c(17, 2, 17, 2, 16, 1, 16, 1)) +
  scale_color_manual(values = c("#f9966d", "#f9966d", "navy", "navy", "#f9966d", "#f9966d", "navy", "navy")) +
  xlab(paste0("PC1 (", percentVar[1], "%)")) +
  ylab(paste0("PC2 (", percentVar[2], "%)")) +
  labs(color = "Group", shape = "Group") +
  theme_bw() + theme(panel.background = element_rect(fill = NA), 
                       plot.title = element_text(hjust = 0.5), 
                       axis.text = element_text(size = 8),
                       axis.title = element_text(size = 8),
                       legend.text = element_text(size=8),
                       legend.title = element_blank(),
                       title = element_text(size=10),
                       aspect.ratio = 1)

ggsave(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/Liver_PCA_Supp_new.pdf",
          plot = g_pcoa, width = 5, height =  3, units = "in", dpi = 300, device = "pdf")


plotting_liver <- function(dds, cat1, cat2) {
  group1 = gsub(" ", "", cat1)
  list1 = strsplit(cat1, split=" ")
  group2 = gsub(" ", "", cat2)
  list2 = strsplit(cat2, split=" ")
  
  note1 = paste0("up in ", list1[[1]][2], " vs ", list2[[1]][2], ", in ", list1[[1]][1], " week old ", list1[[1]][3], " Liver: ")
  note2 = paste0("down in ", list1[[1]][2], " vs ", list2[[1]][2], ", in ", list1[[1]][1], " week old ", list1[[1]][3], " Liver: ")
  
  p_cutoff = 0.05
  fc_cutoff = 1
  
  res <- results(dds, contrast=c("age_fac_sex", group1, group2))
  res_df = as.data.frame(res) %>% mutate(logPvalue = -log10(pvalue))
  
  #Volcano
  up = res_df %>%  dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff) %>% nrow()
  down = res_df %>%  dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff) %>% nrow()
  g <- ggplot(res_df %>% dplyr::filter(padj >= p_cutoff), aes(x = log2FoldChange, y = logPvalue)) + 
      geom_point(size = 1, color = "gray") + 
      geom_point(size = 1, data = res_df %>% dplyr::filter(padj < p_cutoff), 
                 aes(x = log2FoldChange, y = logPvalue),
                 color = "blue") + 
      geom_point(size = 1, data = res_df %>% dplyr::filter(padj < p_cutoff, abs(log2FoldChange) > fc_cutoff),
                 aes(x = log2FoldChange, y = logPvalue),
                 color = "red") + 
      geom_point(size = 1, data = data.frame(log2FoldChange = 0, logPvalue = 0,
                                   category = factor( c("Not significant", "p-value", "p-value and log2 FC"), levels = c("Not significant", "p-value", "p-value and log2 FC"))),
    aes(log2FoldChange, logPvalue, color = category), size = 1, alpha = 0) +
  scale_color_manual(name = "", 
                      values = c("Not significant" = "gray", "p-value" = "blue", "p-value and log2 FC" = "red"), 
                     labels = c("Not significant", "p-value", expression(p-value~and~log[2]~FC))) +
    guides(color = guide_legend(override.aes = list(alpha = 1, size = 1))) +
      labs(x = expression(log[2]~FC), y = expression(-log[10]~(p-value)), 
           #title = paste0("Week ", list1[[1]][1], ", ", list1[[1]][3], "\nVivarium vs Farm")
           ) + 
      theme_bw() + theme(panel.background = element_rect(fill = NA), 
                       plot.title = element_text(hjust = 0.5), 
                       axis.text = element_text(size = 8),
                       axis.title = element_text(size = 8),
                       legend.text = element_text(size=8),
                       legend.title = element_blank(),
                       title = element_text(size=10),
                       aspect.ratio = 1)
    ymax = ggplot_build(g)$layout$panel_scales_y[[1]]$range$range[2]
  return(g)
}    

week4_M <- plotting_liver(liver_dds,"4 Bartlett Male", "4 Farm Male")
week4_F <- plotting_liver(liver_dds,"4 Bartlett Female", "4 Farm Female")
week12_M <- plotting_liver(liver_dds,"12 Bartlett Male", "12 Farm Male")
week12_F <- plotting_liver(liver_dds,"12 Bartlett Female", "12 Farm Female")

p <- week4_M + week12_M +
      week4_F + week12_F +
    plot_layout(guides = "collect", ncol = 2) + 
      plot_annotation(title = "", tag_levels = list(c("B", "C", "D", "E"))) & theme(plot.tag = element_text(size = 12, face = "bold"))
    ggsave(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/Liver_volcano_patchwork.pdf",
            plot = p, width = 8, height = 5, units = "in", dpi = 300, device = "pdf")




p_cutoff = 0.05
fc_cutoff = 1
age_fac_color = c("#7fc97f","#beaed4","#fdc086", "#ffff99")

#up in Bartlett
res <- results(liver_dds, contrast=c("age_fac_sex", "4BartlettMale", "4FarmMale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff)
UP_4BM <- rownames(res_df)

res <- results(liver_dds, contrast=c("age_fac_sex", "12BartlettMale", "12FarmMale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff)
UP_12BM <- rownames(res_df)

res <- results(liver_dds, contrast=c("age_fac_sex", "12BartlettFemale", "12FarmFemale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff)
UP_12BF <- rownames(res_df)

gene_sets_up <- list(
  "Week 4 Male" = UP_4BM,
  "Week 12 Male" = UP_12BM,
  "Week 4 Female" = c(),
  "Week 12 Female" = UP_12BF
  )


#Venn diagram
g_up <- ggvenn(gene_sets_up, fill_color = age_fac_color, set_name_size = 5, text_size = 5, show_percentage = F) + 
  labs(title="Genes Upregulated in Vivarium Mice")
g_up <- g_up + theme(plot.title = element_text(hjust = 0.5, size = 16))


#down in Bartlett
res <- results(liver_dds, contrast=c("age_fac_sex", "4BartlettMale", "4FarmMale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff)
Down_4BM <- rownames(res_df)

res <- results(liver_dds, contrast=c("age_fac_sex", "12BartlettMale", "12FarmMale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff)
Down_12BM <- rownames(res_df)

res <- results(liver_dds, contrast=c("age_fac_sex", "12BartlettFemale", "12FarmFemale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff)
Down_12BF <- rownames(res_df)

gene_sets_down <- list(
  "Week 4 Male" = Down_4BM,
  "Week 12 Male" = Down_12BM,
  "Week 4 Female" = c(),
  "Week 12 Female" = Down_12BF
  )

#Venn diagram
g_down <- ggvenn(gene_sets_down, fill_color = age_fac_color, set_name_size = 5, text_size = 5, show_percentage = F) +
  labs(title="Genes Downregulated in Vivarium Mice")
g_down <- g_down + theme(plot.title = element_text(hjust = 0.5, size = 16))

p <- g_up + g_down +
    plot_layout(guides = "collect", ncol = 2) + 
      plot_annotation(title = "", tag_levels = list(c( "F", "G"))) & theme(plot.tag = element_text(size = 12, face = "bold"))

ggsave(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/Liver_venn_up_supp.pdf",
       plot = g_up, width = 6, height = 4, units = "in", dpi = 300, device = "pdf")
ggsave(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Liver/Liver_venn_down_supp.pdf",
       plot = g_down, width = 6, height = 4, units = "in", dpi = 300, device = "pdf")



```


#Ileum PCA - EXCLUDES OUTLIERS
```{r}
#Outliers: Chlorine0498, Chlorine0502, Chlorine4061

ileumMD$SamplingWeeks <- as.character(ileumMD$SamplingWeeks)
#ileumMD$Sex <- ifelse(ileumMD$Sex=="F", "Female", "Male")
ileumMD$Facility <- ifelse(ileumMD$Facility=="Farm", "Farm", "Vivarium")
ileumMD$age_fac_sex <- paste(ileumMD$SamplingWeeks, ileumMD$Facility, ileumMD$Sex, sep="_")
ileumMD2 <- ileumMD[!(ileumMD$TubeID %in% c("Chlorine0498", "Chlorine0502", "Chlorine4061")), ]

ileum_dds <- DESeqDataSetFromTximport(txi=ileum_txi, colData=ileumMD, design=~Facility + SamplingWeeks + Sex)
ileum_dds <- DESeq(ileum_dds)
ileum_res <- results(ileum_dds)
ileum_dds$age_fac_sex <- factor(paste0(ileumMD$SamplingWeeks, ileumMD$Facility, ileumMD$Sex))
design(ileum_dds) <- ~ age_fac_sex
ileum_dds <- DESeq(ileum_dds)
resultsNames(ileum_dds)
ileum_res <- results(ileum_dds)

coef <- resultsNames(ileum_dds)
coef <- coef[coef != "Intercept"]
clean_names <- sub("^age_fac_sex_", "", coef)
clean_names <- sub("_vs_.*$", "", clean_names)

clean_names


#PCA by Sex
vsdata <- vst(ileum_dds, blind=FALSE)
#plotPCA(vsdata, intgroup="Facility") for quick plotting
p <- plotPCA(vsdata, intgroup="Sex", returnData=TRUE)
percentVar <- round(100 * attr(p, "percentVar"))
p <- p[order(p$Sex),]

mat <- t(assay(vsdata))
meta <- as.data.frame(colData(ileum_dds))
adonis2(
  mat ~ Sex, #can also do each variable individually
  data = meta,
  method = "euclidean",
  permutations = 999, 
  by="margin"
)

#Female = salmon, male = blue
g <- ggplot(p, aes(PC1, PC2, color=Sex)) +
  geom_point(size=3, aes(color=Sex)) +
  scale_color_manual(values = c("salmon", "blue")) +
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  theme_minimal() +
  theme(legend.text = element_text(size = 12), axis.title = element_text(size = 12)) +
  stat_ellipse() +
  ggtitle("Ileum PCA Plot by Sex")
ggsave(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/PCA_by_Sex_Clean.pdf",
          plot = g, width = 7, height =  5, units = "in", dpi = 300, device = "pdf")


#PCA by Age
vsdata <- vst(ileum_dds, blind=FALSE)
#plotPCA(vsdata, intgroup="Facility") for quick plotting
p <- plotPCA(vsdata, intgroup="SamplingWeeks", returnData=TRUE)
percentVar <- round(100 * attr(p, "percentVar"))
p <- p[order(p$SamplingWeeks),]

mat <- t(assay(vsdata))
meta <- as.data.frame(colData(ileum_dds))
adonis2(
  mat ~ SamplingWeeks, #can also do each variable individually
  data = meta,
  method = "euclidean",
  permutations = 999, 
  by="margin"
)

#12weeks = salmon, 4weeks = blue
g <- ggplot(p, aes(PC1, PC2, color=SamplingWeeks)) +
  geom_point(size=3, aes(color=SamplingWeeks)) +
  scale_color_manual(values = c("salmon", "blue")) +
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  theme_minimal() +
  theme(legend.text = element_text(size = 12), axis.title = element_text(size = 12)) +
  stat_ellipse() +
  ggtitle("Ileum PCA Plot by Age")
ggsave(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/PCA_by_Age_Clean.pdf",
          plot = g, width = 7, height =  5, units = "in", dpi = 300, device = "pdf")


#PCA by Facility
vsdata <- vst(ileum_dds, blind=FALSE)
#plotPCA(vsdata, intgroup="Facility") for quick plotting
p <- plotPCA(vsdata, intgroup="Facility", returnData=TRUE)
percentVar <- round(100 * attr(p, "percentVar"))
p <- p[order(p$Facility),]

mat <- t(assay(vsdata))
meta <- as.data.frame(colData(ileum_dds))
adonis2(
  mat ~ Facility, #can also do each variable individually
  data = meta,
  method = "euclidean",
  permutations = 999, 
  by="margin"
)

#Bartlett = salmon, Farm = blue
g <- ggplot(p, aes(PC1, PC2, color=Facility)) +
  geom_point(size=3, aes(color=Facility)) +
  scale_color_manual(values = c("salmon", "blue")) +
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  theme_minimal() +
  theme(legend.text = element_text(size = 12), axis.title = element_text(size = 12)) +
  stat_ellipse() +
  ggtitle("Ileum PCA Plot by Facility")
ggsave(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/PCA_by_Facility_Clean.pdf",
          plot = g, width = 7, height =  5, units = "in", dpi = 300, device = "pdf")





#all groups plotted
vsdata <- vst(ileum_dds, blind=FALSE)
#plotPCA(vsdata, intgroup="Facility") for quick plotting
p <- plotPCA(vsdata, intgroup=c("Sex", "SamplingWeeks", "Facility"), returnData=TRUE)
percentVar <- round(100 * attr(p, "percentVar"))
p$age_sex_fac <- paste0("Week ", ileumMD$SamplingWeeks, ", ", ileumMD$Sex, ", ", ileumMD$Facility)
p <- p[order(p$age_sex_fac),]

mat <- t(assay(vsdata))
meta <- as.data.frame(colData(ileum_dds))
adonis2(
  mat ~ Sex + SamplingWeeks + Facility, #can also do each variable individually
  data = meta,
  method = "euclidean",
  permutations = 999, 
  by="margin"
)

#Female = salmon, male = blue
#4 weeks = circle, 12 weeks = triangle
#solid = bartlett, hollow = farm
g <-ggplot(p, aes(PC1, PC2, color=age_sex_fac)) +
  geom_point(size=3, aes(color=age_sex_fac, shape=age_sex_fac)) +
  scale_shape_manual(values = c(17, 2, 17, 2, 16, 1, 16, 1)) +
  scale_color_manual(values = c("salmon", "salmon", "blue", "blue", "salmon", "salmon", "blue", "blue")) +
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  labs(color = "Group", shape = "Group") +
  theme_minimal() +
  theme(legend.text = element_text(size = 11), axis.title = element_text(size = 12)) #+
  #stat_ellipse() +
  #ggtitle("Ileum PCA Plot")
ggsave(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/Ileum_PCA_supp.pdf",
          plot = g, width = 7, height =  5, units = "in", dpi = 300, device = "pdf")


#ileum metadata, dds, etc. 
#excludes outliers, design = ~Facility + Age
save(ileumMD, ileum_dds, ileum_res, file = "/Users/aiswaryaa/Downloads/CL_rna/Ileum_data.RData")
```


#Load ileum data
```{r}
#load ileum - EXCLUDES OUTLIERS
load("/Users/aiswaryaa/Downloads/CL_rna/Ileum_data.RData")

#load genelist
load("/Users/aiswaryaa/Downloads/CL_rna/genes_txis.RData")

```


#Ileum volcano plots and heatmaps and MA by age and sex
```{r}
#cat order: age, facility, sex
#function to plot volcano and heatmap
plotting_ileum <- function(dds, cat1, cat2) {
  group1 = gsub(" ", "", cat1)
  list1 = strsplit(cat1, split=" ")
  group2 = gsub(" ", "", cat2)
  list2 = strsplit(cat2, split=" ")
  
  note1 = paste0("up in ", list1[[1]][2], " vs ", list2[[1]][2], ", in ", list1[[1]][1], " week old ", list1[[1]][3], " Ileum: ")
  note2 = paste0("down in ", list1[[1]][2], " vs ", list2[[1]][2], ", in ", list1[[1]][1], " week old ", list1[[1]][3], " Ileum: ")
  
  p_cutoff = 0.05
  fc_cutoff = 1
  
  res <- results(dds, contrast=c("age_fac_sex", group1, group2))
  res_df = as.data.frame(res) %>% mutate(logPvalue = -log10(pvalue))
  
  #Volcano
  up = res_df %>%  dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff) %>% nrow()
  down = res_df %>%  dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff) %>% nrow()
  g = ggplot(res_df %>% dplyr::filter(padj >= p_cutoff), aes(x = log2FoldChange, y = logPvalue)) + 
      geom_point(color = "gray") + 
      geom_point(data = res_df %>% dplyr::filter(padj < p_cutoff), 
                 aes(x = log2FoldChange, y = logPvalue),
                 color = "blue") + 
      geom_point(data = res_df %>% dplyr::filter(padj < p_cutoff, abs(log2FoldChange) > fc_cutoff),
                 aes(x = log2FoldChange, y = logPvalue),
                 color = "red") + 
      geom_point(data = data.frame(log2FoldChange = 0, logPvalue = 0,
                                   category = factor( c("Not significant", "p-value", "p-value and log2 FC"), levels = c("Not significant", "p-value", "p-value and log2 FC"))),
    aes(log2FoldChange, logPvalue, color = category), size = 10, alpha = 0) +
  scale_color_manual(name = "", 
                      values = c("Not significant" = "gray", "p-value" = "blue", "p-value and log2 FC" = "red"), 
                     labels = c("Not significant", "p-value", expression(p-value~and~log[2]~FC))) +
    guides(color = guide_legend(override.aes = list(alpha = 1, size = 4))) +
      labs(x = expression(log[2]~FC), y = expression(-log[10]~(p-value)), 
           title = paste0("Week ", list1[[1]][1], "\nVivarium vs Farm, ", list1[[1]][3], ", Ileum")) + 
      theme_bw() + theme(panel.background = element_rect(fill = NA), 
                             plot.title = element_text(hjust = 0.5, size = 16), 
                             text = element_text(color = "black"),
                             axis.text = element_text(size = 14, color = "black"),
                             axis.title = element_text(size = 14),
                             legend.text = element_text(size = 12))
    ymax = ggplot_build(g)$layout$panel_scales_y[[1]]$range$range[2]
    #g = g + geom_text(data = data.frame(log2FoldChange = c(0,0),
    #                             logPvalue = c(ymax*0.95, ymax*0.85),
    #                             note = paste0(c(note1, note2), 
    #                                           c(up, down))),
    #                  aes(x = log2FoldChange, y = logPvalue, label=note), 
    #                  size = 14/.pt, fontface = "bold")
    ggsave(file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/Volcano/", group1 , "_", group2, "_Volcano_Supp.pdf"),
          plot = g, width = 7, height =  5, units = "in", dpi = 300, device = "pdf") 
    
    #MA plot
    g = ggplot(res_df, aes(x = baseMean, y = log2FoldChange)) + 
      geom_point(color = "gray") + 
      geom_point(data = res_df %>% dplyr::filter(padj < p_cutoff), 
                 aes(x = baseMean, y = log2FoldChange),
                 color = "blue") + 
      geom_point(data = res_df %>% dplyr::filter(padj < p_cutoff, abs(log2FoldChange) > fc_cutoff), 
                 aes(x = baseMean, y = log2FoldChange),
                 color = "red") +
      scale_x_log10(breaks = c(1e-1, 1e0, 1e1, 1e2, 1e3, 1e4, 1e5,1e6)) + 
      geom_point(data = data.frame(log2FoldChange = 0, logPvalue = 0,
                                   category = factor( c("Not significant", "Significant (p < 0.05)", "Significant + Large FC\n (p < 0.05 and |LFC| > 1)"), levels = c("Not significant", "Significant (p < 0.05)", "Significant + Large FC\n (p < 0.05 and |LFC| > 1)"))),
    aes(log2FoldChange, logPvalue, color = category), size = 4, alpha = 0) +
  scale_color_manual(name = "Differential Expression", 
                      values = c("Not significant" = "gray", "Significant (p < 0.05)" = "blue", "Significant + Large FC\n (p < 0.05 and |LFC| > 1)" = "red")) +
    guides(color = guide_legend(override.aes = list(alpha = 1, size = 4))) +
      labs(x = "Mean of normalized counts", y = "log2FoldChange (Bartlett/Farm)", title = paste0(list1[[1]][2], " vs ", list2[[1]][2], ", in ", list1[[1]][1], " week old ", list1[[1]][3], " Ileum")) + 
      theme_bw() + theme(panel.background = element_rect(fill = NA), 
                             plot.title = element_text(hjust = 0.5), 
                             text = element_text(face = 'bold', color = "black"),
                             axis.text = element_text(size = 14, color = "black"),
                             axis.title = element_text(size = 14))
    ggsave(file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/MA/", group1 , "_", group2, "_MA_plot_AFSdesign.pdf"),
               plot = g, width = 7, height =  5, units = "in", dpi = 300, device = "pdf")
    
    
    #Heatmap
    sig <- res[!is.na(res$padj) & res$padj < 0.05 & abs(res$log2FoldChange) > 1, ]
    genes <- rownames(sig)
    length(genes)
    vsd <- vst(dds, blind=FALSE)
    mat <- assay(vsd)[genes, ]
    
    group <- colData(dds)$age_fac_sex
    keep <- group %in% c(group1, group2)
    
    mat <- mat[, keep]
    anno <- data.frame(group=group[keep])
    rownames(anno) <- colnames(mat)
    anno <- mutate(anno, group = ifelse(grepl("Bartlett",anno$group), "Bartlett", "Farm"))
    
    mat_scaled <- t(scale(t(mat))) %>% 
      as.data.frame(.) %>% 
      rownames_to_column(., "gene_id") %>% 
      mutate(., mgi_symbol = genelist$mgi_symbol[match(.$gene_id, genelist$ensembl_gene_id)]) %>% 
      column_to_rownames(., "mgi_symbol") %>% 
      .[, -c(1)]
    
    top <- head(order(sig$padj), 50)
    
    pheatmap(mat_scaled[top, ], annotation_col=anno, main = paste0(list1[[1]][1], " week old ", list1[[1]][3], " Ileum"), filename = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/Heatmap/", group1 , "_", group2, "_Heatmap_AFSdesign.pdf")) 

    }

plotting_ileum(ileum_dds,"4 Bartlett Male", "4 Farm Male")
plotting_ileum(ileum_dds,"4 Bartlett Female", "4 Farm Female")
plotting_ileum(ileum_dds,"12 Bartlett Male", "12 Farm Male")
plotting_ileum(ileum_dds,"12 Bartlett Female", "12 Farm Female")
```


#Ileum Volcano plots Combined Sex by age
```{r}
#Week 4 with combined sexes
res <- results(ileum_dds, contrast = list(c("age_fac_sex_4BartlettMale_vs_12BartlettFemale", "age_fac_sex_4BartlettFemale_vs_12BartlettFemale"), c("age_fac_sex_4FarmMale_vs_12BartlettFemale","age_fac_sex_4FarmFemale_vs_12BartlettFemale")),listValues = c(1/2, -1/2))

res_df <- as.data.frame(res) %>% mutate(logPvalue = -log10(pvalue))%>% 
    rownames_to_column(., "gene_id") 
    res_df <- mutate(res_df, mgi_symbol = genelist$mgi_symbol[match(res_df$gene_id, genelist$ensembl_gene_id)]) %>% 
    mutate(res_df, gene_name = mgi_symbol) %>% 
    drop_na() %>% 
    distinct(mgi_symbol, .keep_all = TRUE) %>% 
    column_to_rownames(., "mgi_symbol") %>% 
    .[, -c(1)]
    
    genes_of_interest <- c("Igkv8-21", "C9", "C8a")
    res_df$label <- if_else(res_df$gene_name %in% genes_of_interest, res_df$gene_name, NA_character_)

  up = res_df %>%  dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff) %>% nrow()
  down = res_df %>%  dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff) %>% nrow()
  g = ggplot(res_df %>% dplyr::filter(padj >= p_cutoff), aes(x = log2FoldChange, y = logPvalue)) + 
      geom_point(color = "gray13") + 
      geom_point(data = res_df %>% dplyr::filter(padj < p_cutoff), 
                 aes(x = log2FoldChange, y = logPvalue),
                 color = "blue") + 
      geom_point(data = res_df %>% dplyr::filter(padj < p_cutoff, abs(log2FoldChange) > fc_cutoff),
                 aes(x = log2FoldChange, y = logPvalue),
                 color = "red") + 
      geom_point(data = data.frame(log2FoldChange = 0, logPvalue = 0,
                                   category = factor( c("Not significant", "p < 0.05", "p < 0.05 and |log2 FC| > 1"), levels = c("Not significant", "p < 0.05", "p < 0.05 and |log2 FC| > 1"))),
    aes(log2FoldChange, logPvalue, color = category), size = 10, alpha = 0) +
  scale_color_manual(name = "", 
                      values = c("Not significant" = "gray13", "p < 0.05" = "blue", "p < 0.05 and |log2 FC| > 1" = "red"), 
                     labels = c("Not significant", "p < 0.05", "p < 0.05 and")) +
    guides(color = guide_legend(override.aes = list(alpha = 1, size = 4))) +
      labs(x = expression(log[2]~FC(Vivarium/Farm)), y = expression(-log[10]~(p-value)), 
           title = paste0("Week 4, Ileum")) + 
      theme(plot.title = element_text(hjust = 0.5, size = 16), 
                             text = element_text(color = "black"),
                             axis.text = element_text(size = 16, color = "black"),
                             axis.title = element_text(size = 16),
                             legend.text = element_text(size = 16)) +
    geom_label_repel(
    data = res_df %>% dplyr::filter(!is.na(label)),
    aes(x = log2FoldChange, y = logPvalue, label = label),
    size = 5,
    show.legend = FALSE,
    min.segment.length = 0
  )
    ymax = ggplot_build(g)$layout$panel_scales_y[[1]]$range$range[2]
g

ggsave(file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/RUMP_Week4_Ileum_Volcano.pdf"), plot = g, width = 7, height =  5, units = "in", dpi = 300, device = "pdf")


    
#Week 12 with combined sexes
    res <- results(ileum_dds, contrast = list(c("age_fac_sex_12BartlettMale_vs_12BartlettFemale"), c("age_fac_sex_12FarmMale_vs_12BartlettFemale","age_fac_sex_12FarmFemale_vs_12BartlettFemale")),listValues = c(1/2, -1/2))
    
    res_df <- as.data.frame(res) %>% mutate(logPvalue = -log10(pvalue))%>% 
    rownames_to_column(., "gene_id") 
    res_df <- mutate(res_df, mgi_symbol = genelist$mgi_symbol[match(res_df$gene_id, genelist$ensembl_gene_id)]) %>% 
    mutate(res_df, gene_name = mgi_symbol) %>% 
    drop_na() %>% 
    distinct(mgi_symbol, .keep_all = TRUE) %>% 
    column_to_rownames(., "mgi_symbol") %>% 
    .[, -c(1)]
    
    genes_of_interest <- c("Ccl21d", "Spon2", "Apol9b", "Nfil3", "Ccl20", "Ccl17", "Socs3", "Igkv9-124", "Ighv5-16")
    res_df$label <- if_else(res_df$gene_name %in% genes_of_interest, res_df$gene_name, NA_character_)

  up = res_df %>%  dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff) %>% nrow()
  down = res_df %>%  dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff) %>% nrow()
  g = ggplot(res_df %>% dplyr::filter(padj >= p_cutoff), aes(x = log2FoldChange, y = logPvalue)) + 
      geom_point(color = "gray13") + 
      geom_point(data = res_df %>% dplyr::filter(padj < p_cutoff), 
                 aes(x = log2FoldChange, y = logPvalue),
                 color = "blue") + 
      geom_point(data = res_df %>% dplyr::filter(padj < p_cutoff, abs(log2FoldChange) > fc_cutoff),
                 aes(x = log2FoldChange, y = logPvalue),
                 color = "red") + 
      geom_point(data = data.frame(log2FoldChange = 0, logPvalue = 0,
                                   category = factor( c("Not significant", "p < 0.05", "p < 0.05 and |log2 FC| > 1"), levels = c("Not significant", "p < 0.05", "p < 0.05 and |log2 FC| > 1"))),
    aes(log2FoldChange, logPvalue, color = category), size = 10, alpha = 0) +
  scale_color_manual(name = "", 
                      values = c("Not significant" = "gray13", "p < 0.05" = "blue", "p < 0.05 and |log2 FC| > 1" = "red"), 
                     labels = c("Not significant", "p < 0.05", "p < 0.05 and")) +
    guides(color = guide_legend(override.aes = list(alpha = 1, size = 4))) +
      labs(x = expression(log[2]~FC(Vivarium/Farm)), y = expression(-log[10]~(p-value)), 
           title = paste0("Week 12, Ileum")) + 
      theme(plot.title = element_text(hjust = 0.5, size = 16), 
                             text = element_text(color = "black"),
                             axis.text = element_text(size = 16, color = "black"),
                             axis.title = element_text(size = 16),
                             legend.text = element_text(size = 16)) +
    geom_label_repel(
    data = res_df %>% dplyr::filter(!is.na(label)),
    aes(x = log2FoldChange, y = logPvalue, label = label),
    size = 5,
    show.legend = FALSE,
    min.segment.length = 0
  )
    ymax = ggplot_build(g)$layout$panel_scales_y[[1]]$range$range[2]
g

  ggsave(file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/RUMP_Week12_Ileum_Volcano.pdf"), plot = g, width = 7, height =  5, units = "in", dpi = 300, device = "pdf")

up_sig_res <- res_df %>% dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff)
up_named <- as.data.frame(up_sig_res) %>% mutate(logPvalue = -log10(pvalue))%>% 
    rownames_to_column(., "gene_id") %>% 
    drop_na() %>% 
    distinct(gene_name, .keep_all = TRUE) %>% 
    column_to_rownames(., "gene_id") %>% 
    .[, -c(1)]
  

up_sig_res <- res_df %>% dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff)
named <- as.data.frame(up_sig_res) %>% mutate(logPvalue = -log10(pvalue))%>% 
    rownames_to_column(., "gene_id") 

down_sig_res <- res_df %>% dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff)
named <- as.data.frame(down_sig_res) %>% mutate(logPvalue = -log10(pvalue))%>% 
    rownames_to_column(., "gene_id") 


down_sig_res <- res_df %>% dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff)
down_named <- as.data.frame(down_sig_res) %>% mutate(logPvalue = -log10(pvalue))%>% 
    rownames_to_column(., "gene_id") %>% 
    drop_na() %>% 
    distinct(gene_name, .keep_all = TRUE) %>% 
    column_to_rownames(., "gene_id") %>% 
    .[, -c(1)]

```


#Ileum Heatmap Main Figure
```{r}

res <- results(ileum_dds, contrast=c("age_fac_sex", "4BartlettMale", "4FarmMale"))
male4_sig <- res[!is.na(res$padj) & res$padj < 0.05 & abs(res$log2FoldChange) > 2, ]

res <- results(ileum_dds, contrast=c("age_fac_sex", "12BartlettMale", "12FarmMale"))
male12_sig <- res[!is.na(res$padj) & res$padj < 0.05 & abs(res$log2FoldChange) > 2, ]

res <- results(ileum_dds, contrast=c("age_fac_sex", "4BartlettFemale", "4FarmFemale"))
female4_sig <- res[!is.na(res$padj) & res$padj < 0.05 & abs(res$log2FoldChange) > 2, ]

res <- results(ileum_dds, contrast=c("age_fac_sex", "12BartlettFemale", "12FarmFemale"))
female12_sig <- res[!is.na(res$padj) & res$padj < 0.05 & abs(res$log2FoldChange) > 2, ]

genes <- c(rownames(male4_sig), rownames(male12_sig), rownames(female4_sig),rownames(female12_sig))
#lengths(genes) = 75

vsd <- vst(ileum_dds, blind=FALSE)
mat <- assay(vsd)[genes, ]

group <- colData(ileum_dds)$age_fac_sex
keep <- group %in% c("4BartlettMale", "4FarmMale", "12BartlettMale", "12FarmMale", "4BartlettFemale", "4FarmFemale", "12BartlettFemale", "12FarmFemale")

mat <- mat[, keep]
anno <- data.frame(group=group[keep])
rownames(anno) <- colnames(mat)
anno <- mutate(anno, Sex = ifelse(grepl("Female",anno$group), "Female", "Male"),
               Age = ifelse(grepl("4",anno$group), "4 weeks", "12 weeks"),
               Facility = ifelse(grepl("Bartlett",anno$group), "Vivarium", "Farm")) %>% 
  .[, -1]

mat_scaled <- t(scale(t(mat))) %>% 
  as.data.frame(.) %>% 
  rownames_to_column(., "gene_id") %>% 
  mutate(., mgi_symbol = genelist$mgi_symbol[match(.$gene_id, genelist$ensembl_gene_id)]) %>% 
  drop_na() %>% 
  column_to_rownames(., "mgi_symbol") %>% 
  .[, -c(1)]

#New original heatmap
ileumMD$Facility <- ifelse(ileumMD$Facility=="Farm", "Farm", "Vivarium")
temp_meta <- ileumMD[order(ileumMD$Facility), ]
temp_meta <- mutate(temp_meta, 
                    SamplingWeeks = ifelse(grepl("4",temp_meta$SamplingWeeks), "4 weeks", "12 weeks"),
                    Facility = factor(Facility, levels = c("Vivarium", "Farm"))) %>%
  arrange(Facility)
mat <- mat[, temp_meta$TubeID]

col_fun = colorRamp2(c(-4, 0, 4), c("royalblue", "lemonchiffon", "red1"))
#highlights cells with p < .1
#col_fun = colorRamp2(c(-4, -1.3, 0, 1.3, 4), c("blue", "white", "white", "white", "red"))

col_anno <- HeatmapAnnotation(
  Age = temp_meta$SamplingWeeks,
  Facility = temp_meta$Facility,
  Sex = temp_meta$Sex,
  annotation_name_gp = gpar(fontsize = 10),
  col = list(
    Facility = c("Vivarium" = "dodgerblue", "Farm" = "forestgreen"),
    Age = c("4 weeks" = "gold", "12 weeks" = "darkorange"),
    Sex = c("Male" = "darkmagenta", "Female" = "firebrick2")
  )
)

#circadian_genes = c("Per1", "Per2", "Per3", "Dbp", "Ciart", "Bmal1", "Nr1d1", "Nr1d2")
mat_scaled <- t(scale(t(mat))) %>% 
  as.data.frame(.) %>% 
  rownames_to_column(., "gene_id") %>% 
  mutate(., mgi_symbol = genelist$mgi_symbol[match(.$gene_id, genelist$ensembl_gene_id)]) %>% 
  drop_na() %>% 
  column_to_rownames(., "mgi_symbol") %>% 
  .[, -c(1)] %>% 
  as.matrix() #%>% .[rownames(.) %in% circadian_genes, ]

p <- Heatmap(
  mat_scaled,
  #col = col_fun,
  name = "Z-score",
  top_annotation = col_anno,
  cluster_columns = TRUE,      
  show_column_dend = TRUE,      
  show_row_dend = TRUE,
  column_title = NULL,
  show_column_names = FALSE,
  row_names_gp = gpar(fontsize = 8),
  column_title_side = "bottom"
)

pdf(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/Ileum_Heatmap_main2.pdf", width = 7, height = 8)
p
dev.off()

#gives data according to draw figure
p2 = draw(p)
row_indices = row_order(p2)
col_indices = column_order(p2)
ordered_mat = mat_scaled[row_indices, col_indices]

#Supervised by facility
col_anno <- HeatmapAnnotation(
  Facility = temp_meta$Facility,
  Age = temp_meta$SamplingWeeks,
  Sex = temp_meta$Sex,
  annotation_name_gp = gpar(fontsize = 10),
  col = list(
    Facility = c("Vivarium" = "#252525", "Farm" = "#1b9e77"),
    Age = c("4 weeks" = "#f9966d", "12 weeks" = "navy"),
    Sex = c("Male" = "#2b8cbe", "Female" = "#f1726d")
  )
)

p <- Heatmap(
  mat_scaled,
  width = ncol(mat_scaled) * unit(2, "mm"), 
  height = nrow(mat_scaled) * unit(3, "mm"),
  name = "Z-score",
  top_annotation = col_anno,
  column_split = temp_meta$Facility,
  cluster_columns = TRUE,
  cluster_column_slices = FALSE,
  #cluster_columns = TRUE,       # unsupervised clustering within each block
  #column_split = temp_meta$Facility,        # separate blocks by facility
  show_column_dend = TRUE,      # dendrograms for each block
  show_row_dend = TRUE,
  column_title = "Differentially Expressed Genes in Ileum",
  show_column_names = FALSE,
  row_names_gp = gpar(fontsize = 8),
  column_title_side = "top",
  column_title_gp = gpar(fontface = "bold")
)

cairo_pdf(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/Ileum_Heatmap_facility_supervised.pdf", width = 8, height = 10, fallback_resolution = 300)
p
dev.off()


#Supervised by sex
col_anno <- HeatmapAnnotation(
  Sex = temp_meta$Sex,
  Age = temp_meta$SamplingWeeks,
  Facility = temp_meta$Facility,
  annotation_name_gp = gpar(fontsize = 10),
  col = list(
    Sex = c("Male" = "darkmagenta", "Female" = "firebrick2"),
    Facility = c("Vivarium" = "dodgerblue", "Farm" = "forestgreen"),
    Age = c("4 weeks" = "gold", "12 weeks" = "darkorange")
  )
)

p <- Heatmap(
  mat_scaled,
  name = "Z-score",
  top_annotation = col_anno,
  cluster_columns = TRUE,       # unsupervised clustering within each block
  column_split = temp_meta$Sex,        # separate blocks by facility
  show_column_dend = TRUE,      # dendrograms for each block
  show_row_dend = TRUE,
  column_title = NULL,
  show_column_names = FALSE,
  row_names_gp = gpar(fontsize = 8),
  column_title_side = "bottom"
)

pdf(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/Ileum_Heatmap_sex_supervised.pdf", width = 8, height = 6)
p
dev.off()


#Supervised by age
col_anno <- HeatmapAnnotation(
  Age = temp_meta$SamplingWeeks,
  Sex = temp_meta$Sex,
  Facility = temp_meta$Facility,
  annotation_name_gp = gpar(fontsize = 10),
  col = list(
    Sex = c("Male" = "darkmagenta", "Female" = "firebrick2"),
    Facility = c("Vivarium" = "dodgerblue", "Farm" = "forestgreen"),
    Age = c("4 weeks" = "gold", "12 weeks" = "darkorange")
  )
)

p <- Heatmap(
  mat_scaled,
  name = "Z-score",
  top_annotation = col_anno,
  cluster_columns = TRUE,       # unsupervised clustering within each block
  column_split = temp_meta$SamplingWeeks,        # separate blocks by facility
  show_column_dend = TRUE,      # dendrograms for each block
  show_row_dend = TRUE,
  column_title = NULL,
  show_column_names = FALSE,
  row_names_gp = gpar(fontsize = 8),
  column_title_side = "bottom"
)

pdf(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/Ileum_Heatmap_age_supervised.pdf", width = 8, height = 6)
p
dev.off()

```


#Ileum UpSet plot and venn diagram 
```{r}

p_cutoff = 0.05
fc_cutoff = 1
age_fac_color = c("#7fc97f","#beaed4","#fdc086", "#ffff99")

#up in Bartlett
res <- results(ileum_dds, contrast=c("age_fac_sex", "4BartlettMale", "4FarmMale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff)
UP_4BM <- rownames(res_df)

res <- results(ileum_dds, contrast=c("age_fac_sex", "12BartlettMale", "12FarmMale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff)
UP_12BM <- rownames(res_df)

res <- results(ileum_dds, contrast=c("age_fac_sex", "4BartlettFemale", "4FarmFemale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff)
UP_4BF <- rownames(res_df)

res <- results(ileum_dds, contrast=c("age_fac_sex", "12BartlettFemale", "12FarmFemale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff)
UP_12BF <- rownames(res_df)

gene_sets <- list(
  "Week 4 Male" = UP_4BM,
  "Week 12 Male" = UP_12BM,
  "Week 4 Female" = UP_4BF,
  "Week 12 Female" = UP_12BF
  )

#run this in R console for it to work
plotObject <- UpSetR::upset(fromList(gene_sets), text.scale = c(1,1,1,1,2,2))
pdf(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/Upset_Venn_diagram/Upset_up_in_Bartlett_AFSdesign.pdf", width = 8, height = 6)
plotObject
dev.off()

#Venn diagram
g <- ggvenn(gene_sets, fill_color = age_fac_color, set_name_size = 5, text_size = 5, show_percentage = F) +
  labs(title="Genes Upregulated in Vivarium Mice")
g <- g + theme(plot.title = element_text(hjust = 0.5, size = 14))
ggsave(file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/Upset_Venn_diagram/Venn_up_in_Bartlett_supp.pdf"),
         plot = g, width = 7, height =  5, units = "in", dpi = 300, device = "pdf")



#down in Bartlett
res <- results(ileum_dds, contrast=c("age_fac_sex", "4BartlettMale", "4FarmMale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff)
Down_4BM <- rownames(res_df)

res <- results(ileum_dds, contrast=c("age_fac_sex", "12BartlettMale", "12FarmMale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff)
Down_12BM <- rownames(res_df)

res <- results(ileum_dds, contrast=c("age_fac_sex", "4BartlettFemale", "4FarmFemale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff)
Down_4BF <- rownames(res_df)

res <- results(ileum_dds, contrast=c("age_fac_sex", "12BartlettFemale", "12FarmFemale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff)
Down_12BF <- rownames(res_df)

gene_sets <- list(
  "Week 4 Male" = Down_4BM,
  "Week 12 Male" = Down_12BM,
  "Week 4 Female" = Down_4BF,
  "Week 12 Female" = Down_12BF
  )

#run this in R console for it to work
plotObject <- UpSetR::upset(fromList(gene_sets), text.scale = c(1,1,1,1,2,2))
pdf(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/Upset_Venn_diagram/Upset_down_in_Bartlett_AFSdesign.pdf", width = 8, height = 6)
plotObject
dev.off()

#Venn diagram
g <- ggvenn(gene_sets, fill_color = age_fac_color, set_name_size = 5, text_size = 5, show_percentage = F) +
  labs(title="Genes Downregulated in Vivarium Mice")
g <- g + theme(plot.title = element_text(hjust = 0.5, size = 14))
ggsave(file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/Upset_Venn_diagram/Venn_down_in_Bartlett_supp.pdf"),
         plot = g, width = 7, height =  5, units = "in", dpi = 300, device = "pdf")

```


#Ileum GAGE/pathview pathways
```{r}
data(kegg.sets.mm)
data(sigmet.idx.mm)
kegg.sets.mm = kegg.sets.mm[sigmet.idx.mm]

ileum_pathways <- function(dds, cat1, cat2){
    group1 = gsub(" ", "", cat1)
    list1 = strsplit(cat1, split=" ")
    group2 = gsub(" ", "", cat2)
    list2 = strsplit(cat2, split=" ")
    pathway = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/pathways/", group1, "_vs_", group2)
    setwd(pathway)
    
    res_df <- results(dds, contrast=c("age_fac_sex", group1, group2)) %>% 
    as.data.frame() %>% rownames_to_column(var = "ensembl_gene_id") %>% 
    left_join(genelist %>% dplyr::select(ensembl_gene_id, entrezgene_id) %>% unique(), by = "ensembl_gene_id") %>% 
    dplyr::filter(!is.na(entrezgene_id))
    deseq2.fc = res_df$log2FoldChange
    names(deseq2.fc) = res_df$entrezgene_id
    fc.kegg.p = gage(deseq2.fc, gsets = kegg.sets.mm, ref =NULL, samp = NULL)
    res = rbind(fc.kegg.p$greater %>% as.data.frame() %>% 
      dplyr::filter(q.val < 0.1),
      fc.kegg.p$less %>% as.data.frame() %>% 
      dplyr::filter(q.val < 0.1)) %>% 
      dplyr::filter(!is.na(q.val))    
    res$path_id = str_sub(rownames(res),start = 1, end = 8)
    res = res %>% rownames_to_column(var = "path")
    out.suffix = paste0(list1[[1]][2], " vs ", list2[[1]][2], " in ", list1[[1]][1], " week old ", list1[[1]][3], " Ileum")
    write.csv(res, file = paste0(out.suffix, ".csv"))
    if(nrow(res)>0){
        pv.out.list = sapply(res$path_id, 
                             function(pid) {
                               pathview(gene.data = deseq2.fc, pathway.id = pid, kegg.native = T, 
                                        same.layer = F, species = "mmu",out.suffix=out.suffix)})} 
  
}
ileum_pathways(ileum_dds, "4 Bartlett Male", "4 Farm Male")
ileum_pathways(ileum_dds,"4 Bartlett Female", "4 Farm Female")
ileum_pathways(ileum_dds,"12 Bartlett Male", "12 Farm Male")
ileum_pathways(ileum_dds,"12 Bartlett Female", "12 Farm Female")

getwd()
```


#Ileum preparation for GSEA
```{r}
#msigdbr_species()

vsd <- vst(ileum_dds, blind=FALSE)
b= "Ileum"
for (grp in c("4_Bartlett_Female", "4_Bartlett_Male", "12_Bartlett_Female", "12_Bartlett_Male", "4_Farm_Female", "4_Farm_Male", "12_Farm_Female", "12_Farm_Male" )){
    mat = assay(vsd) %>% as.matrix()
    #colnames(mat) = colData(vsd)$TubeID
    mat = as.data.frame(mat) %>% rownames_to_column(var = "ensembl_gene_id") %>% 
      left_join(genelist %>% dplyr::select(ensembl_gene_id, entrezgene_id) %>% unique() , 
                by = "ensembl_gene_id") %>% 
      dplyr::filter(!is.na(entrezgene_id)) %>%
      dplyr::select(-ensembl_gene_id) %>% column_to_rownames(var = "entrezgene_id") %>% as.matrix()    
    meta_c = ileumMD[str_detect(ileumMD$age_fac_sex, grp),]
    mat = mat[,meta_c$TubeID]
    out <- mat %>% as.data.frame() %>% 
      rownames_to_column(var = "NAME") %>% 
      mutate(Description = "") %>% 
      dplyr::select(NAME, Description, everything())
    
    con <- file(paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/for_gsea/", b, "_", grp, "_for_gsea.txt"), open="wt")
    #writeLines(paste0("# test"), con)
    write.table(out, con, sep = "\t", quote = F, row.names = F)   
    close(con)
    
    meta_c <- meta_c %>% mutate(code = ifelse(Facility == "Bartlett", 0, 1))
    con <- file(paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/for_gsea/", b, "_", grp, "_for_gsea.cls"), open="wt")
    writeLines(paste(nrow(meta_c),  meta_c$code %>% unique() %>% length(), 1,sep = "\t"), con)
    writeLines(paste("#", "Bartlett", "Farm",sep = "\t"), con)
    writeLines(paste0(as.character(meta_c$code), collapse = "\t"), con)
    close(con)
  }

```

#Ileum GSEA, rank by LFC
```{r}

b = "Ileum"
res_kegg <- list()
res_go <- list()
res_gsea <- list()

#sex_time_color = c("#7fc97f","#beaed4","#fdc086", "#ffff99")

#HARDCODE EACH GROUP  
cat1="12 Bartlett Female"
cat2="12 Farm Female"
  group1 = gsub(" ", "", cat1)
  list1 = strsplit(cat1, split=" ") #age fac sex
  group2 = gsub(" ", "", cat2)
  list2 = strsplit(cat2, split=" ")
  grp = paste0(list1[[1]][1], "_", list1[[1]][3])

  
  res_df <-  results(ileum_dds, contrast = c("age_fac_sex", group1, group2)) %>% 
      as.data.frame() %>% rownames_to_column(var = "ensembl_gene_id") %>% 
      left_join(genelist %>% dplyr::select(ensembl_gene_id, entrezgene_id) %>% unique(), 
                by = "ensembl_gene_id") %>% 
      dplyr::filter(!is.na(entrezgene_id), !is.na(log2FoldChange))
    
    d <- res_df %>% dplyr::select(entrezgene_id, log2FoldChange) %>% 
      arrange(desc(log2FoldChange))
    geneList <- d[,2]
    names(geneList) <- as.character(d[,1])
    
    gsea_go <- gseGO(geneList = geneList,
                     OrgDb = org.Mm.eg.db,
                     ont = "ALL",
                     keyType = "ENTREZID",
                     pAdjustMethod = "BH",
                     pvalueCutoff  = 0.05,
                     eps = 0)
    res <- gsea_go@result %>% mutate(Tissue = b, grp = grp)
    res_go <- rbind(res_go, res)
    res_gsea[["go"]][[b]][[grp]] <- gsea_go
    
    gsea_kegg <- gseKEGG(geneList = geneList,
                         organism = "mmu",
                         keyType = "ncbi-geneid",
                         pAdjustMethod = "BH",
                         pvalueCutoff  = 0.05,
                         eps = 0)
    
    res <- gsea_kegg@result %>% mutate(Tissue = b, grp = grp)
    res_kegg <- rbind(res_kegg, res)
    res_gsea[["kegg"]][[b]][[grp]] <-  gsea_kegg    


res_go_all <- dplyr::bind_rows(res_go) %>% mutate(Tissue = "Ileum_Bartlett_vs_Farm")
write.table(res_go_all, file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/clusterProfiler/gsea_go_all_LFC.tsv"),
          quote = F, row.names = F,sep = "\t")
as_table <- res_go_all[, c(2,3,5,6,8,13,14)]
write.table(as_table, file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/clusterProfiler/as_table_gsea_go_all_LFC.tsv"),
          quote = F, row.names = F,sep = "\t")

res_kegg_all <- dplyr::bind_rows(res_kegg) %>% mutate(Tissue = "Ileum_Bartlett_vs_Farm")
write.table(res_kegg_all, file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/clusterProfiler/gsea_kegg_all_LFC.tsv"),
          quote = F, row.names = F,sep = "\t")
as_table <- res_kegg_all[, c(1,2,4,5,7,12,13)]
write.table(as_table, file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/clusterProfiler/as_table_gsea_kegg_all_LFC.tsv"),
          quote = F, row.names = F,sep = "\t")

save(res_gsea, res_go_all, res_kegg_all, file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/gsea.Rdata")


```


#Ileum GSEA, rank by LFC*(-log(p-value))
```{r}

b = "Ileum"
res_kegg <- list()
res_go <- list()
res_gsea <- list()

#sex_time_color = c("#7fc97f","#beaed4","#fdc086", "#ffff99")

#HARDCODE EACH GROUP  
cat1="12 Bartlett Female"
cat2="12 Farm Female"
  group1 = gsub(" ", "", cat1)
  list1 = strsplit(cat1, split=" ") #age fac sex
  group2 = gsub(" ", "", cat2)
  list2 = strsplit(cat2, split=" ")
  grp = paste0(list1[[1]][1], "_", list1[[1]][3])

  
  res_df <-  results(ileum_dds, contrast = c("age_fac_sex", group1, group2)) %>% 
      as.data.frame() %>% rownames_to_column(var = "ensembl_gene_id") %>% 
      left_join(genelist %>% dplyr::select(ensembl_gene_id, entrezgene_id) %>% unique(), 
                by = "ensembl_gene_id") %>% 
      dplyr::filter(!is.na(entrezgene_id), !is.na(log2FoldChange), !is.na(pvalue))
    
    d <- res_df %>% dplyr::select(entrezgene_id, log2FoldChange) %>% mutate(rank_val = (res_df$log2FoldChange * -log(res_df$pvalue))) %>% 
      arrange(desc(rank_val))
    geneList <- d[,3]
    names(geneList) <- as.character(d[,1])
    
    gsea_go <- gseGO(geneList = geneList,
                     OrgDb = org.Mm.eg.db,
                     ont = "ALL",
                     keyType = "ENTREZID",
                     pAdjustMethod = "BH",
                     pvalueCutoff  = 0.05,
                     eps = 0)
    res <- gsea_go@result %>% mutate(Tissue = b, grp = grp)
    res_go <- rbind(res_go, res)
    res_gsea[["go"]][[b]][[grp]] <- gsea_go
    
    gsea_kegg <- gseKEGG(geneList = geneList,
                         organism = "mmu",
                         keyType = "ncbi-geneid",
                         pAdjustMethod = "BH",
                         pvalueCutoff  = 0.05,
                         eps = 0)
    
    res <- gsea_kegg@result %>% mutate(Tissue = b, grp = grp)
    res_kegg <- rbind(res_kegg, res)
    res_gsea[["kegg"]][[b]][[grp]] <-  gsea_kegg    


res_go_all <- dplyr::bind_rows(res_go) %>% mutate(Tissue = "Ileum_Bartlett_vs_Farm")
write.table(res_go_all, file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/clusterProfiler/gsea_go_all_LFCxlogPV.tsv"),
          quote = F, row.names = F,sep = "\t")
as_table <- res_go_all[, c(2,3,5,6,8,13,14)]
write.table(as_table, file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/clusterProfiler/as_table_gsea_go_all_LFCxlogPV.tsv"),
          quote = F, row.names = F,sep = "\t")

res_kegg_all <- dplyr::bind_rows(res_kegg) %>% mutate(Tissue = "Ileum_Bartlett_vs_Farm")
write.table(res_kegg_all, file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/clusterProfiler/gsea_kegg_all_LFCxlogPV.tsv"),
          quote = F, row.names = F,sep = "\t")
as_table <- res_kegg_all[, c(1,2,4,5,7,12,13)]
write.table(as_table, file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/clusterProfiler/as_table_gsea_kegg_all_LFCxlogPV.tsv"),
          quote = F, row.names = F,sep = "\t")

```


#Ileum Pathview/Gene Exploration
```{r}
#4 weeks female: mmu04610, mmu00380, mmu04750, mmu04820, mmu04630, mmu00590
#12 weeks female: mmu04662, mmu03015, mmu04610, mmu00190
#folder: pathview_GSEA_KEGG_sig_lfc

data(kegg.sets.mm)
data(sigmet.idx.mm)
kegg.sets.mm = kegg.sets.mm[sigmet.idx.mm]

#4 weeks female
setwd("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/pathview_GSEA_KEGG_sig_lfc/4BartlettFemale_vs_4FarmFemale")
res_df <- results(ileum_dds, contrast=c("age_fac_sex", "4BartlettFemale", "4FarmFemale")) %>% 
    as.data.frame() %>% rownames_to_column(var = "ensembl_gene_id") %>% 
    left_join(genelist %>% dplyr::select(ensembl_gene_id, entrezgene_id) %>% unique(), by = "ensembl_gene_id") %>% 
    dplyr::filter(!is.na(entrezgene_id))
    deseq2.fc = res_df$log2FoldChange
    names(deseq2.fc) = res_df$entrezgene_id
    pv.out.list = sapply(c("mmu04610", "mmu00380", "mmu04750", "mmu04820", "mmu04630", "mmu00590"),
                             function(pid) {
                               pathview(gene.data = deseq2.fc, pathway.id = pid, kegg.native = T, 
                                        same.layer = F, species = "mmu",out.suffix="Bartlett vs Farm in 4 week old Female Ileum GSEA KEGG")}) 
  
#12 weeks female
setwd("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/pathview_GSEA_KEGG_sig_lfc/12BartlettFemale_vs_12FarmFemale")
res_df <- results(ileum_dds, contrast=c("age_fac_sex", "12BartlettFemale", "12FarmFemale")) %>% 
    as.data.frame() %>% rownames_to_column(var = "ensembl_gene_id") %>% 
    left_join(genelist %>% dplyr::select(ensembl_gene_id, entrezgene_id) %>% unique(), by = "ensembl_gene_id") %>% 
    dplyr::filter(!is.na(entrezgene_id))
    deseq2.fc = res_df$log2FoldChange
    names(deseq2.fc) = res_df$entrezgene_id
    pv.out.list = sapply(c("mmu04662", "mmu03015", "mmu04610", "mmu00190"),
                             function(pid) {
                               pathview(gene.data = deseq2.fc, pathway.id = pid, kegg.native = T, 
                                        same.layer = F, species = "mmu",out.suffix="Bartlett vs Farm in 12 week old Female Ileum GSEA KEGG")}) 

    
    

#plotting gene vst counts
vsd <- vst(ileum_dds, blind=FALSE)

#4 weeks
res <- results(ileum_dds, contrast=c("age_fac_sex", "4BartlettFemale", "4FarmFemale"))
sig <- res[!is.na(res$padj) & res$padj < 0.05 & abs(res$log2FoldChange) > 1, ]
female_4weeks_genes <- rownames(sig)
#12 weeks
res <- results(ileum_dds, contrast=c("age_fac_sex", "12BartlettFemale", "12FarmFemale"))
sig <- res[!is.na(res$padj) & res$padj < 0.05 & abs(res$log2FoldChange) > 1, ]
female_12weeks_genes <- rownames(sig)

genes <- c(female_4weeks_genes, female_12weeks_genes) %>% unique()
mat <- assay(vsd)[tgenes, ]

group <- colData(ileum_dds)$age_fac_sex
keep <- group %in% c("4BartlettFemale", "4FarmFemale", "12BartlettFemale", "12FarmFemale")

mat <- mat[, keep] %>% 
  as.data.frame(.) %>%
  rownames_to_column(., "gene_id") %>% 
  mutate(., mgi_symbol = genelist$mgi_symbol[match(.$gene_id, genelist$ensembl_gene_id)]) %>% 
  #dplyr::filter(mgi_symbol %in% genes) %>%  #only if using explicit gene list below
  column_to_rownames(., "mgi_symbol") %>% 
  .[, -c(1)]
mat <- as.data.frame(t(mat))

#main CCC genes of interest(red dots in at least one group): C8a, C8b, C9, Cfhr4, Cpn2, Ighv5-9, Igkv16-104, Vtn

genes <- colnames(mat)
#CCC visually inspected genes
#genes <- c("Fga", "Serpinc1", "F7", "F10", "F5", "F2", "F13b", "F9", "F11", "F12", "Serpine1", "Serping1", "Cpb2", "Plg", "Kng1", "Bdkrb1", "C4bp", "C4b", "C2", "Masp1", "C3", "Hc", "C6", "Clu", "Cfd", "Cfhr1", "Vsig4", "Itgax")
#genes <- c("C1qa", "A2m") #these go from up to downregulated in Bartlett

#B-cell visually inspected genes
#genes <- c("Lilrb4b", "Cd22", "Cd72", "Cd79a", "Cd79b", "Bank1", "Blk", "Fcgr2b", "Cd19", "Cr2", "Ptpn6", "Syk", "Btk", "Vav1", "Inpp5d", "Pik3ap1", "Pik3ca", "Blnk", "Grb2", "Plcg2", "Prkcb", "Akt1", "Rasgrp3", "Rac1", "Ppp3ca", "Malt1", "Card11", "Ikbkb", "Nfkbia", "Nfatc1")
for (gene in (genes)){
  plot_data <- data.frame(counts = mat[, gene], row.names = rownames(mat)) %>% 
  rownames_to_column("TubeID") %>% 
  left_join(select(ileumMD, TubeID, age_fac_sex), by = "TubeID") %>% 
  mutate(Group = gsub("_", " ", .$age_fac_sex))
  
  plot_data$Group <- factor(plot_data$Group, levels = c("4 Bartlett Female", "4 Farm Female", "12 Bartlett Female", "12 Farm Female"))
  
  my_comparisons <- list(
  c("4 Bartlett Female", "4 Farm Female"),
  c("4 Bartlett Female", "12 Bartlett Female"),
  c("4 Bartlett Female", "12 Farm Female"),
  c("4 Farm Female", "12 Bartlett Female"),
  c("4 Farm Female", "12 Farm Female"),
  c("12 Bartlett Female", "12 Farm Female")
)
  
  #ggplot2 boxplot
  #ggplot(plot_data, aes(x = age_fac_sex, y = counts, fill = age_fac_sex)) +
  #  geom_boxplot() +
  #  geom_jitter(width = 0.1, alpha = 0.5) +
  #  labs(title = paste("VST Normalized Counts for", "Gnmt"),
  #       x = "Group",
  #       y = "VST Normalized Counts") +
  #  theme_minimal()
  
  #ggpubr boxplot, with p-adj, else: p.adjust.method = "none"
  g <- ggboxplot(plot_data, x = "Group", y = "counts", 
          color = "Group", add = "jitter",
          xlab = "Group", ylab = "VST Normalized Counts", title = paste0("VST Normalized Counts for ", gene)) +
  geom_pwc(method = "wilcox.test", label = "p.signif", hide.ns = TRUE, p.adjust.method = "none")
    #stat_compare_means(comparisons = my_comparisons, label = "p.signif", hide.ns=T)
  ggsave(file = paste0("/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/VST_boxplots_visual_inspection/", gene, "_Bartlett_vs_Farm_Females.pdf"),
         plot = g, width = 7, height =  5, units = "in", dpi = 300, device = "pdf")
}


```

#Ileum Supp figure
#PCA, volcano(patchwork), venn diagram
```{r}

ileumMD$SamplingWeeks <- as.character(ileumMD$SamplingWeeks)
#ileumMD$Sex <- ifelse(ileumMD$Sex=="F", "Female", "Male")
ileumMD$Facility <- ifelse(ileumMD$Facility=="Farm", "Farm", "Vivarium")
ileumMD$age_fac_sex <- paste(ileumMD$SamplingWeeks, ileumMD$Facility, ileumMD$Sex, sep="_")

#PCA
vsdata <- vst(ileum_dds, blind=FALSE)
p <- plotPCA(vsdata, intgroup=c("Sex", "SamplingWeeks", "Facility"), returnData=TRUE)
percentVar <- round(100 * attr(p, "percentVar"), 2)
p$age_sex_fac <- paste0("Week ", ileumMD$SamplingWeeks, ", ", ileumMD$Sex, ", ", ileumMD$Facility)
p <- p[order(p$age_sex_fac),]

mat <- t(assay(vsdata))
meta <- as.data.frame(colData(ileum_dds))

#Female = salmon, male = blue
#4 weeks = circle, 12 weeks = triangle
#solid = bartlett, hollow = farm
g_pcoa <- ggplot(p, aes(PC1, PC2, color=age_sex_fac)) +
  geom_point(size=3, aes(color=age_sex_fac, shape=age_sex_fac)) +
  scale_shape_manual(values = c(17, 2, 17, 2, 16, 1, 16, 1)) +
  #Age = c("4 weeks" = "#f9966d", "12 weeks" = "navy")
  scale_color_manual(values = c("#f9966d", "#f9966d", "navy", "navy", "#f9966d", "#f9966d", "navy", "navy")) +
  xlab(paste0("PC1 (", percentVar[1], "%)")) +
  ylab(paste0("PC2 (", percentVar[2], "%)")) +
  labs(color = "Group", shape = "Group") +
  theme_bw() + theme(panel.background = element_rect(fill = NA), 
                       plot.title = element_text(hjust = 0.5), 
                       axis.text = element_text(size = 8),
                       axis.title = element_text(size = 8),
                       legend.text = element_text(size=8),
                       legend.title = element_blank(),
                       title = element_text(size=10),
                       aspect.ratio = 1)

ggsave(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/Ileum_PCA_Supp_new.pdf",
          plot = g_pcoa, width = 5, height = 3, units = "in", dpi = 300, device = "pdf")


plotting_ileum <- function(dds, cat1, cat2) {
  group1 = gsub(" ", "", cat1)
  list1 = strsplit(cat1, split=" ")
  group2 = gsub(" ", "", cat2)
  list2 = strsplit(cat2, split=" ")
  
  note1 = paste0("up in ", list1[[1]][2], " vs ", list2[[1]][2], ", in ", list1[[1]][1], " week old ", list1[[1]][3], " Ileum: ")
  note2 = paste0("down in ", list1[[1]][2], " vs ", list2[[1]][2], ", in ", list1[[1]][1], " week old ", list1[[1]][3], " Ileum: ")
  
  p_cutoff = 0.05
  fc_cutoff = 1
  
  res <- results(dds, contrast=c("age_fac_sex", group1, group2))
  res_df = as.data.frame(res) %>% mutate(logPvalue = -log10(pvalue))
  
  #Volcano
  up = res_df %>%  dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff) %>% nrow()
  down = res_df %>%  dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff) %>% nrow()
  g <- ggplot(res_df %>% dplyr::filter(padj >= p_cutoff), aes(x = log2FoldChange, y = logPvalue)) + 
      geom_point(size = 1, color = "gray") + 
      geom_point(size = 1, data = res_df %>% dplyr::filter(padj < p_cutoff), 
                 aes(x = log2FoldChange, y = logPvalue),
                 color = "blue") + 
      geom_point(size = 1, data = res_df %>% dplyr::filter(padj < p_cutoff, abs(log2FoldChange) > fc_cutoff),
                 aes(x = log2FoldChange, y = logPvalue),
                 color = "red") + 
      geom_point(size = 1, data = data.frame(log2FoldChange = 0, logPvalue = 0,
                                   category = factor( c("Not significant", "p-value", "p-value and log2 FC"), levels = c("Not significant", "p-value", "p-value and log2 FC"))),
    aes(log2FoldChange, logPvalue, color = category), size = 1, alpha = 0) +
  scale_color_manual(name = "", 
                      values = c("Not significant" = "gray", "p-value" = "blue", "p-value and log2 FC" = "red"), 
                     labels = c("Not significant", "p-value", expression(p-value~and~log[2]~FC))) +
    guides(color = guide_legend(override.aes = list(alpha = 1, size = 1))) +
      labs(x = expression(log[2]~FC), y = expression(-log[10]~(p-value)), 
           #title = paste0("Week ", list1[[1]][1], ", ", list1[[1]][3], "\nVivarium vs Farm")
           ) + 
      theme_bw() + theme(panel.background = element_rect(fill = NA), 
                       plot.title = element_text(hjust = 0.5), 
                       axis.text = element_text(size = 8),
                       axis.title = element_text(size = 8),
                       legend.text = element_text(size=8),
                       legend.title = element_blank(),
                       title = element_text(size=10),
                       aspect.ratio = 1)
    ymax = ggplot_build(g)$layout$panel_scales_y[[1]]$range$range[2]
  return(g)
}    

week4_M <- plotting_ileum(ileum_dds,"4 Bartlett Male", "4 Farm Male")
week4_F <- plotting_ileum(ileum_dds,"4 Bartlett Female", "4 Farm Female")
week12_M <- plotting_ileum(ileum_dds,"12 Bartlett Male", "12 Farm Male")
week12_F <- plotting_ileum(ileum_dds,"12 Bartlett Female", "12 Farm Female")

p <- week4_M + week12_M +
      week4_F + week12_F +
    plot_layout(guides = "collect", ncol = 2) + 
      plot_annotation(title = "", tag_levels = list(c("B", "C", "D", "E"))) & theme(plot.tag = element_text(size = 12, face = "bold"))
    ggsave(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/Ileum_volcano_patchwork.pdf",
            plot = p, width = 8, height = 5, units = "in", dpi = 300, device = "pdf")




p_cutoff = 0.05
fc_cutoff = 1
age_fac_color = c("#7fc97f","#beaed4","#fdc086", "#ffff99")

#up in Bartlett
res <- results(ileum_dds, contrast=c("age_fac_sex", "4BartlettMale", "4FarmMale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff)
UP_4BM <- rownames(res_df)

res <- results(ileum_dds, contrast=c("age_fac_sex", "12BartlettMale", "12FarmMale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff)
UP_12BM <- rownames(res_df)

res <- results(ileum_dds, contrast=c("age_fac_sex", "12BartlettFemale", "12FarmFemale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange > fc_cutoff)
UP_12BF <- rownames(res_df)

gene_sets_up <- list(
  "Week 4 Male" = UP_4BM,
  "Week 12 Male" = UP_12BM,
  "Week 4 Female" = c(),
  "Week 12 Female" = UP_12BF
  )


#Venn diagram
g_up <- ggvenn(gene_sets_up, fill_color = age_fac_color, set_name_size = 5, text_size = 5, show_percentage = F) + 
  labs(title="Genes Upregulated in Vivarium Mice")
g_up <- g_up + theme(plot.title = element_text(hjust = 0.5, size = 16))


#down in Bartlett
res <- results(ileum_dds, contrast=c("age_fac_sex", "4BartlettMale", "4FarmMale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff)
Down_4BM <- rownames(res_df)

res <- results(ileum_dds, contrast=c("age_fac_sex", "12BartlettMale", "12FarmMale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff)
Down_12BM <- rownames(res_df)

res <- results(ileum_dds, contrast=c("age_fac_sex", "12BartlettFemale", "12FarmFemale"))
res_df = as.data.frame(res) %>% dplyr::filter(padj < p_cutoff, log2FoldChange < -fc_cutoff)
Down_12BF <- rownames(res_df)

gene_sets_down <- list(
  "Week 4 Male" = Down_4BM,
  "Week 12 Male" = Down_12BM,
  "Week 4 Female" = c(),
  "Week 12 Female" = Down_12BF
  )

#Venn diagram
g_down <- ggvenn(gene_sets_down, fill_color = age_fac_color, set_name_size = 5, text_size = 5, show_percentage = F) +
  labs(title="Genes Downregulated in Vivarium Mice")
g_down <- g_down + theme(plot.title = element_text(hjust = 0.5, size = 16))

p <- g_up + g_down +
    plot_layout(guides = "collect", ncol = 2) + 
      plot_annotation(title = "", tag_levels = list(c( "F", "G"))) & theme(plot.tag = element_text(size = 12, face = "bold"))

ggsave(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/Ileum_venn_supp.pdf",
       plot = p, width = 10, height = 4, units = "in", dpi = 300, device = "pdf")
ggsave(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/Ileum_venn_up_supp.pdf",
       plot = g_up, width = 6, height = 4, units = "in", dpi = 300, device = "pdf")
ggsave(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results/Ileum/Ileum_venn_down_supp.pdf",
       plot = g_down, width = 6, height = 4, units = "in", dpi = 300, device = "pdf")



```

#Gene functions in upregulated heatmap clusters(liver and ileum)
```{r}
#Liver
dat <- read_excel("/Users/aiswaryaa/Downloads/CL_rna/Liver Main Figure Gene Function Table.xlsx") %>% as.data.frame()

colnames(dat)[1] <- "Group"
dat_l <- dat %>%
  pivot_longer(
    cols = -Group,
    names_to = "GeneFunction",
    values_to = "Count"
  ) %>%
  replace_na(list(Count = 0)) 

dat_l$Group <- factor(dat_l$Group, c("Farm Adults", 
                                     "Vivarium Adults",
                                     "Females", 
                                     "Adult Males"))
dat_l$GeneFunction <- factor(dat_l$GeneFunction, levels = colnames(dat)[-1])

#Ileum
dat <- read_excel("/Users/aiswaryaa/Downloads/CL_rna/Ileum Main Figure Gene Function Table.xlsx") %>% as.data.frame()

colnames(dat)[1] <- "Group"
dat_i <- dat %>%
  pivot_longer(
    cols = -Group,
    names_to = "GeneFunction",
    values_to = "Count"
  ) %>%
  replace_na(list(Count = 0))
dat_i$GeneFunction <- factor(dat_i$GeneFunction, levels = colnames(dat)[-1])

#facet by tissue
dat_l$Tissue <- "Liver"
dat_i$Tissue <- "Ileum"
dat_c <- rbind(dat_l, dat_i)
dat_c$Tissue <- factor(dat_c$Tissue, c("Liver", "Ileum"))

col_pal <- c("#3366CC", "#DC3912", "#FF9900", "#109618", "#990099", "#0099C6", "#DD4477", "darkgreen", "#D55E00", "#004586", "#997511")
names(col_pal) <- unique(dat_c$GeneFunction) %>% as.factor()

g <- ggplot(dat_c, aes(x = Group, y = Count, fill = GeneFunction)) +
  geom_bar(stat = "identity") +
  labs(x = "", y = "Number of Upregulated Genes") +
  facet_grid(~Tissue, scales = "free_x", space = "free_x") +
  theme_bw() + theme(panel.background = element_rect(fill = NA), 
                   plot.title = element_text(hjust = 0.5), 
                   axis.text = element_text(size = 10),
                   axis.title = element_text(size = 10),
                   legend.text = element_text(size=8),
                   legend.title = element_text(size=8, face = "bold"),
                   title = element_text(size=10),
                   strip.text = element_text(color = "black", face = "bold", size = 10),
                   strip.background = element_blank(),
                   axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = col_pal, name = "Gene Function")

ggsave(file = "/Users/aiswaryaa/Downloads/CL_rna/DESeq_Results//Main_Heatmap_gene_functions_panelC.pdf", plot = g, width = 6, height = 5, units = "in", dpi = 300, device = "pdf")

```

